---
title: "Data Prep & Cleaning"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(sf)
```

```{r making not in function}
'%!in%' <- Negate('%in%')
```

### 1. Loading data

```{r load data}
## List of permitted CAFOs in North Carolina - 2024, NC Department of Enviro Quality
cafo_df <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/CAFOs NC.csv")

## ID incidence (per 100,000) by NC county - 2023, NCDHHS
countyid_df <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/County Chart_data.csv")

## NC county area and number of hospital beds - 2023, Office of State Budget and Management and 2025 NC DHHS
cb_df <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC area and beds.csv")

## percent in poverty by county - 2022, Office of State Budget and Management
poverty_df <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/perc_in_poverty.csv")

## percent foreign-born - 2023, US Census, American Community Survey
foreign_df <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/ACS 2023 5 year.csv")
```

### 2. Swine CAFOs

We have multiple entries for each CAFO type because of the different kinds of permits. But we only want swine, and we don't want multiple entries per company at one location. Let's organize that.

```{r CAFO reformatting}
table(cafo_df$Regulated.Operation)
swine_df <- cafo_df[cafo_df$Regulated.Operation=="Swine",]
swine_df <- swine_df %>%
  distinct(Permit.Number, Location.Lat.Num, Location.Long.Num, .keep_all = TRUE)
saveRDS(swine_df, "swine_df.RDS")
```

### 3. Diseases

We're only going to look at:

-   TB
-   other respiratory (grouped)
-   vector-borne diseases (grouped)
-   food-/water-borne diseases (grouped)
-   healthcare-associated and AMR (grouped)

```{r ID reformatting}
## selecting 2023
id2023_df <- countyid_df[countyid_df$Year.of.Year==2023 & countyid_df$Measure.Names=="County Incidence",]
table(id2023_df$Disease)

## find the unique diseases
unique_IDs <- unique(id2023_df$Disease)

## get rid of diseases that don't appear in any county
dis_num <- vector(length=100)
empties <- vector()
for(i in 1:length(unique_IDs)){
  loop_df <- id2023_df[id2023_df$Disease==unique_IDs[i],]
  dis_num <- sum(loop_df$Measure.Values, na.rm=TRUE)
  if(dis_num==0){
    empties <- c(empties, unique_IDs[i])
  }
  print(i)
}

id2023_df <- id2023_df[id2023_df$Disease %!in% empties,]
  
# update the unique disease list
unique_IDs <- unique(id2023_df$Disease)

## categories
disease_type <- rep(NA, length(unique_IDs))
ID_types <- data.frame(unique_IDs, disease_type)
colnames(ID_types) <- c("Disease", "disease_type")

# TB
ID_types$disease_type[ID_types$Disease=="TB"] <- "TB"

# other respiratory
ID_types$disease_type <- ifelse(
  ID_types$Disease=="Pertussis" | ID_types$Disease=="Mumps" | 
    ID_types$Disease=="Meningococcal invasive disease" | 
    ID_types$Disease=="Meningitis, pneumococcal" | 
    ID_types$Disease=="Legionellosis" |
    ID_types$Disease=="Haemophilus influenzae, invasive disease"|
    ID_types$Disease=="Influenza, adult death" | 
    ID_types$Disease=="Influenza, pediatric death", 
  "resp", ID_types$disease_type
)

# vector-borne
ID_types$disease_type <- ifelse(
  ID_types$Disease=="Anaplasmosis" | ID_types$Disease=="Chikungunya" |
    ID_types$Disease=="Dengue" | ID_types$Disease=="Ehrlichiosis, chaffeensis" |
    ID_types$Disease=="Ehrlichiosis, unspecified" |
    ID_types$Disease=="Encephalitis, arboviral, LAC" |
    ID_types$Disease=="Encephalitis, arboviral, WNV" |
    ID_types$Disease=="Lyme Disease" | ID_types$Disease=="Malaria" |
    ID_types$Disease=="Spotted Fever Rickettsiosis" |
    ID_types$Disease=="Tularemia", 
  "vec", ID_types$disease_type
)

# food-/water-borne
ID_types$disease_type <- ifelse(
  ID_types$Disease=="Botulism - foodborne/wound" | 
    ID_types$Disease=="Campylobacteriosis" | 
    ID_types$Disease=="Cryptosporidiosis" | ID_types$Disease=="Cyclosporiasis" |
    ID_types$Disease=="E. coli - shiga toxin producing" |
    ID_types$Disease=="Foodborne dz- C. perfringens" |
    ID_types$Disease=="Foodborne dz- other/unknown" |
    ID_types$Disease=="Foodborne dz- Staphylococcal" |
    ID_types$Disease=="Listeriosis" | ID_types$Disease=="Salmonellosis" |
    ID_types$Disease=="Shigellosis" | 
    ID_types$Disease=="Hemolytic Uremic Syndrome", 
  "fw", ID_types$disease_type
)

# HAI and AMR
ID_types$disease_type <- ifelse(
  ID_types$Disease=="Candida auris" |
    ID_types$Disease=="Carbapenem-resistant Enterobacterales" |
    ID_types$Disease=="Staph aureus, reduced suscept. to vancomycin" |
    ID_types$Disease=="Streptococcal invasive infection, Group A", 
  "ha", ID_types$disease_type
)

## drop everything that isn't categorized
ID_types <- ID_types[!is.na(ID_types$disease_type),]

## join back in, and drop everything that isn't categorized
id2023_df <- left_join(id2023_df, ID_types, by="Disease")
id2023_df <- id2023_df[!is.na(id2023_df$disease_type),]

## get raw number of cases for each disease
id2023_df$n_case <- NA
id2023_df$n_case <- id2023_df$Measure.Values*id2023_df$County.Pop.Adjusted/100000

## now get total numbers per disease category per county
id2023_df$tot_n <- NA
tot_df <- id2023_df %>%
  group_by(County, disease_type) %>%
  summarise(dis_cat_tot=sum(n_case))

## now, IRs
county_pop <- vector()
counties <- unique(id2023_df$County)
for(i in 1:length(counties)){
  county_pop[i] <- max(id2023_df$County.Pop.Adjusted[id2023_df$County==counties[i]])
}
county_pop <- data.frame(counties, county_pop)
colnames(county_pop) <- c("County", "county_pop")

tot_df <- left_join(tot_df, county_pop, by="County")
tot_df$c_IR <- NA
tot_df$c_IR <- tot_df$dis_cat_tot/tot_df$county_pop*100000
```

Using *tot_df* we'll make the final data table that we'll use for analyses.
```{r making df}
df <- tot_df[,-3]
df <- pivot_wider(df, names_from="disease_type", values_from="c_IR")
df$resp <- round(df$resp, digits=1)
```

### 4. Healthcare access and county size

Unfortunately, we needed to hardcode this. We added it into the csv of county area.
```{r hospital beds}
df <- left_join(df, cb_df, by="County")
colnames(df)[8] <- "land_area"
colnames(df)[9] <- "hosp_beds"

df$beds_pc <- df$hosp_beds/df$county_pop

df <- df[,-9] # drop hospital bed number

```

### 5. Poverty

We will add the percent of county living in poverty.
```{r poverty}
df <- left_join(df, poverty_df, by="County")

## hard-code New Hanover County
df$Percent.in.Poverty[df$County=="New Hanover"] <- 12.1
```


### 6. Combining CAFO and ID dataframes

Finally, we want density of CAFOs per county (CAFOs per square mile). 
```{r CAFO density}
cbc_table <- as.data.frame(table(swine_df$County.Name))
colnames(cbc_table) <- c("County", "cafo_num")

df <- left_join(df, cbc_table)
df$cafo_num <- ifelse(is.na(df$cafo_num), 0, df$cafo_num)

df$cafo_dens <- df$cafo_num/df$land_area
```

### 7. Population density
```{r popdens}
df$pop_dens <- df$county_pop/df$land_area
```

### 8. Foreign-born
```{r foreign-born}
foreign_df <- foreign_df[116,]
keep_foreign <- vector()
for(i in 1:100){
  keep_foreign[i] <- 6+((i-1)*4)
}
foreign_df <- foreign_df[,keep_foreign]
foreign_df <- pivot_longer(foreign_df, cols=1:100, names_to="County")
colnames(foreign_df) <- c("County", "foreign")
foreign_df$County <- str_remove(foreign_df$County, ".County..North.Carolina..Estimate")
foreign_df$foreign <- str_remove(foreign_df$foreign, ",")
foreign_df$foreign <- as.numeric(foreign_df$foreign)

df <- left_join(df, foreign_df, by="County")
df$foreign <- str_remove(df$foreign, ",")
df$foreign <- as.numeric(df$foreign)

df <- df %>%
  mutate(f_perc=foreign/county_pop*100)

## hard code New Hanover County
df$f_perc[65] <- 6.07
```

### 9. Save
```{r saving dfs}
saveRDS(swine_df, "swine_df.RDS")
saveRDS(df, "df.RDS")
```

