---
title: "Data Prep & Cleaning"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(sf)
library(tidycensus)
```

```{r tidycensus setup}
## DO NOT RUN if you already have a Census API key loaded
census_api_key("7e09c574348b580868c23ac0304cdea01e65413c", install = TRUE)
```

```{r making not in function}
'%!in%' <- Negate('%in%')
```

### 1. Loading data

Infectious disease incidence data were downloaded from NCDHHS (https://epi.dph.ncdhhs.gov/cd/figures.html). Data were downloaded one year at a time due to R import issues with multiple-year spreadsheets. 

These data also have the adjusted population sizes of each county.

```{r NC infectious diseases}
id_2017 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2017.csv")

id_2018 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2018.csv")

id_2019 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2019.csv")

id_2020 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2020.csv")

id_2021 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2021.csv")

id_2022 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2022.csv")

id_2023 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/Infectious diseases/NC_ID_2023.csv")
```

2017 and 2020 CAFO data come from personal correspondence with Dr. Christine Lawson; these spreadsheets were originally publicly available on the NC DEQ website but were replaced by the 2024 spreadsheet (https://www.deq.nc.gov/about/divisions/water-resources/permitting/animal-feeding-operations/animal-facility-map).

```{r CAFOs in NC}
cafo_2017 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/CAFOs/CAFO 2017.csv")

cafo_2020 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/CAFOs/CAFO 2020.csv")

cafo_2024 <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/CAFOs/CAFO 2024.csv")
```

Land area data for each county as of 1 July 2024 were obtained from the Office of State Budget and Management (https://www.osbm.nc.gov/facts-figures/population-demographics/state-demographer/municipal-population-estimates/municipal-and-non-municipal-land-area).

```{r land area}
land_area <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/State info/NC land area.csv")
```

Percent insured by county in North Carolina were obtained from ACS 5-year estimates for 2017-2023. This will be used to get percent uninsured estimates and serves as a proxy for healthcare access. We use 5-year estimates because not all of our counties are big enough for 1-year estimates.

```{r percent insured}
unins_2017 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2017
)

unins_2018 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2018
)

unins_2019 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2019
)

unins_2020 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2020
)

unins_2021 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2021
)

unins_2022 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2022
)

unins_2023 <- get_acs(
  geography = "county",
  variables = c(perc_unins = "S2701_C05_001"),
  state = "NC",
  year = 2023
)
```

The poverty rate by county in North Carolina were obtained from ACS 5-year estimates for 2017-2023.

```{r poverty}
pov_2017 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2017
)

pov_2018 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2018
)

pov_2019 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2019
)

pov_2020 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2020
)

pov_2021 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2021
)

pov_2022 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2022
)

pov_2023 <- get_acs(
  geography = "county",
  variables = c(perc_pov = "S1701_C03_001"),
  state = "NC",
  year = 2023
)
```

The percent foreign-born by county in  North Carolina were obtained from the North Carolina Office of State Budget and Management.

```{r foreign born}
foreign_born <- read.csv("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/Data/State info/foreign-born-population-and-citizenship.csv")
```


### 2. Diseases

We're only going to look at:

-   TB
-   other respiratory (grouped)

```{r diseases 2017}
## fix variable types
id_2017$County.Pop.Adjusted <- as.numeric(id_2017$County.Pop.Adjusted)
id_2017$County.Cases <- as.numeric(id_2017$County.Cases)
id_2017$County.Incidence <- as.numeric(id_2017$County.Incidence)

## set aside population sizes
pop_2017 <- id_2017[seq(1, nrow(id_2017), by=9),]
pop_2017 <- pop_2017 %>%
  select(County, County.Pop.Adjusted)

## specify TB and ORDs
id_2017$type[id_2017$Disease=="TB"] <- "TB"
id_2017$type <- ifelse(
    id_2017$Disease=="Pertussis" | id_2017$Disease=="Mumps" | 
    id_2017$Disease=="Meningococcal invasive disease" | 
    id_2017$Disease=="Legionellosis" |
    id_2017$Disease=="Haemophilus influenzae, invasive disease"|
    id_2017$Disease=="Influenza, adult death" | 
    id_2017$Disease=="Influenza, pediatric death" | 
    id_2017$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2017$type
)
id_2017 <- id_2017[!is.na(id_2017$type),]

## get pooled IRs for the ORDs
id_2017$n_case <- id_2017$County.Incidence*id_2017$County.Pop.Adjusted/100000

resp_2017 <- id_2017 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2017 <- left_join(resp_2017, pop_2017, by = "County")

resp_2017$p_ir <- resp_2017$cat_n/resp_2017$County.Pop.Adjusted*100000
```

```{r diseases 2018}
## fix variable types
id_2018$County.Pop.Adjusted <- as.numeric(id_2018$County.Pop.Adjusted)
id_2018$County.Cases <- as.numeric(id_2018$County.Cases)
id_2018$County.Incidence <- as.numeric(id_2018$County.Incidence)

## specify TB and ORDs
id_2018$type[id_2018$Disease=="TB"] <- "TB"
id_2018$type <- ifelse(
    id_2018$Disease=="Pertussis" | id_2018$Disease=="Mumps" | 
    id_2018$Disease=="Meningococcal invasive disease" | 
    id_2018$Disease=="Legionellosis" |
    id_2018$Disease=="Haemophilus influenzae, invasive disease"|
    id_2018$Disease=="Influenza, adult death" | 
    id_2018$Disease=="Influenza, pediatric death" | 
    id_2018$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2018$type
)
id_2018 <- id_2018[!is.na(id_2018$type),]

## set aside population sizes
pop_2018 <- id_2018[seq(1, nrow(id_2018), by=9),]
pop_2018 <- pop_2018 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2018$n_case <- id_2018$County.Incidence*id_2018$County.Pop.Adjusted/100000

resp_2018 <- id_2018 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2018 <- left_join(resp_2018, pop_2018, by = "County")

resp_2018$p_ir <- resp_2018$cat_n/resp_2018$County.Pop.Adjusted*100000
```

```{r diseases 2019}
## fix variable types
id_2019$County.Pop.Adjusted <- as.numeric(id_2019$County.Pop.Adjusted)
id_2019$County.Cases <- as.numeric(id_2019$County.Cases)
id_2019$County.Incidence <- as.numeric(id_2019$County.Incidence)

## specify TB and ORDs
id_2019$type[id_2019$Disease=="TB"] <- "TB"
id_2019$type <- ifelse(
    id_2019$Disease=="Pertussis" | id_2019$Disease=="Mumps" | 
    id_2019$Disease=="Meningococcal invasive disease" | 
    id_2019$Disease=="Legionellosis" |
    id_2019$Disease=="Haemophilus influenzae, invasive disease"|
    id_2019$Disease=="Influenza, adult death" | 
    id_2019$Disease=="Influenza, pediatric death" | 
    id_2019$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2019$type
)
id_2019 <- id_2019[!is.na(id_2019$type),]

## set aside population sizes
pop_2019 <- id_2019[seq(1, nrow(id_2019), by=9),]
pop_2019 <- pop_2019 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2019$n_case <- id_2019$County.Incidence*id_2019$County.Pop.Adjusted/100000

resp_2019 <- id_2019 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2019 <- left_join(resp_2019, pop_2019, by = "County")

resp_2019$p_ir <- resp_2019$cat_n/resp_2019$County.Pop.Adjusted*100000
```

```{r diseases 2020}
## fix variable types
id_2020$County.Pop.Adjusted <- as.numeric(id_2020$County.Pop.Adjusted)
id_2020$County.Cases <- as.numeric(id_2020$County.Cases)
id_2020$County.Incidence <- as.numeric(id_2020$County.Incidence)

## specify TB and ORDs
id_2020$type[id_2020$Disease=="TB"] <- "TB"
id_2020$type <- ifelse(
    id_2020$Disease=="Pertussis" | id_2020$Disease=="Mumps" | 
    id_2020$Disease=="Meningococcal invasive disease" | 
    id_2020$Disease=="Legionellosis" |
    id_2020$Disease=="Haemophilus influenzae, invasive disease"|
    id_2020$Disease=="Influenza, adult death" | 
    id_2020$Disease=="Influenza, pediatric death" | 
    id_2020$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2020$type
)
id_2020 <- id_2020[!is.na(id_2020$type),]

## set aside population sizes
pop_2020 <- id_2020[seq(1, nrow(id_2020), by=9),]
pop_2020 <- pop_2020 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2020$n_case <- id_2020$County.Incidence*id_2020$County.Pop.Adjusted/100000

resp_2020 <- id_2020 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2020 <- left_join(resp_2020, pop_2020, by = "County")

resp_2020$p_ir <- resp_2020$cat_n/resp_2020$County.Pop.Adjusted*100000
```

```{r diseases 2021}
## fix variable types
id_2021$County.Pop.Adjusted <- as.numeric(id_2021$County.Pop.Adjusted)
id_2021$County.Cases <- as.numeric(id_2021$County.Cases)
id_2021$County.Incidence <- as.numeric(id_2021$County.Incidence)

## specify TB and ORDs
id_2021$type[id_2021$Disease=="TB"] <- "TB"
id_2021$type <- ifelse(
    id_2021$Disease=="Pertussis" | id_2021$Disease=="Mumps" | 
    id_2021$Disease=="Meningococcal invasive disease" | 
    id_2021$Disease=="Legionellosis" |
    id_2021$Disease=="Haemophilus influenzae, invasive disease"|
    id_2021$Disease=="Influenza, adult death" | 
    id_2021$Disease=="Influenza, pediatric death" | 
    id_2021$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2021$type
)
id_2021 <- id_2021[!is.na(id_2021$type),]

## set aside population sizes
pop_2021 <- id_2021[seq(1, nrow(id_2021), by=9),]
pop_2021 <- pop_2021 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2021$n_case <- id_2021$County.Incidence*id_2021$County.Pop.Adjusted/100000

resp_2021 <- id_2021 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2021 <- left_join(resp_2021, pop_2021, by = "County")

resp_2021$p_ir <- resp_2021$cat_n/resp_2021$County.Pop.Adjusted*100000
```

```{r diseases 2022}
## fix variable types
id_2022$County.Pop.Adjusted <- as.numeric(id_2022$County.Pop.Adjusted)
id_2022$County.Cases <- as.numeric(id_2022$County.Cases)
id_2022$County.Incidence <- as.numeric(id_2022$County.Incidence)

## specify TB and ORDs
id_2022$type[id_2022$Disease=="TB"] <- "TB"
id_2022$type <- ifelse(
    id_2022$Disease=="Pertussis" | id_2022$Disease=="Mumps" | 
    id_2022$Disease=="Meningococcal invasive disease" | 
    id_2022$Disease=="Legionellosis" |
    id_2022$Disease=="Haemophilus influenzae, invasive disease"|
    id_2022$Disease=="Influenza, adult death" | 
    id_2022$Disease=="Influenza, pediatric death" | 
    id_2022$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2022$type
)
id_2022 <- id_2022[!is.na(id_2022$type),]

## set aside population sizes
pop_2022 <- id_2022[seq(1, nrow(id_2022), by=9),]
pop_2022 <- pop_2022 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2022$n_case <- id_2022$County.Incidence*id_2022$County.Pop.Adjusted/100000

resp_2022 <- id_2022 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2022 <- left_join(resp_2022, pop_2022, by = "County")

resp_2022$p_ir <- resp_2022$cat_n/resp_2022$County.Pop.Adjusted*100000
```

```{r diseases 2023}
## fix variable types
id_2023$County.Pop.Adjusted <- as.numeric(id_2023$County.Pop.Adjusted)
id_2023$County.Cases <- as.numeric(id_2023$County.Cases)
id_2023$County.Incidence <- as.numeric(id_2023$County.Incidence)

## specify TB and ORDs
id_2023$type[id_2023$Disease=="TB"] <- "TB"
id_2023$type <- ifelse(
    id_2023$Disease=="Pertussis" | id_2023$Disease=="Mumps" | 
    id_2023$Disease=="Meningococcal invasive disease" | 
    id_2023$Disease=="Legionellosis" |
    id_2023$Disease=="Haemophilus influenzae, invasive disease"|
    id_2023$Disease=="Influenza, adult death" | 
    id_2023$Disease=="Influenza, pediatric death" | 
    id_2023$Disease=="Influenza, NOVEL virus infection",
    "ORD", id_2023$type
)
id_2023 <- id_2023[!is.na(id_2023$type),]

## set aside population sizes
pop_2023 <- id_2023[seq(1, nrow(id_2023), by=9),]
pop_2023 <- pop_2023 %>%
  select(County, County.Pop.Adjusted)

## get pooled IRs for the ORDs
id_2023$n_case <- id_2023$County.Incidence*id_2023$County.Pop.Adjusted/100000

resp_2023 <- id_2023 %>%
  group_by(County, type) %>%
  summarize(cat_n = sum(n_case))
resp_2023 <- left_join(resp_2023, pop_2023, by = "County")

resp_2023$p_ir <- resp_2023$cat_n/resp_2023$County.Pop.Adjusted*100000
```

### 3. Swine CAFOs

We have multiple entries for each CAFO type because of the different kinds of permits. But we only want swine, and we don't want multiple entries per company at one location. Let's organize that.

```{r CAFO reformatting}
table(cafo_df$Regulated.Operation)
swine_df <- cafo_df[cafo_df$Regulated.Operation=="Swine",]
swine_df <- swine_df %>%
  distinct(Permit.Number, Location.Lat.Num, Location.Long.Num, .keep_all = TRUE)
saveRDS(swine_df, "swine_df.RDS")
```

### 4. Healthcare access and county size

Unfortunately, we needed to hardcode this. We added it into the csv of county area.
```{r hospital beds}
df <- left_join(df, cb_df, by="County")
colnames(df)[8] <- "land_area"
colnames(df)[9] <- "hosp_beds"

df$beds_pc <- df$hosp_beds/df$county_pop

df <- df[,-9] # drop hospital bed number

```

### 5. Poverty

We will add the percent of county living in poverty.
```{r poverty}
colnames(poverty_df)[1] <- "County"
poverty_df$County <- str_remove(poverty_df$County, " County")
df <- left_join(df, poverty_df, by="County")
df <- df %>%
  select(-Area.Type, -Year, -Variable) %>%
  rename(Percent.in.Poverty = Value)
```


### 6. Combining CAFO and ID dataframes

Finally, we want density of CAFOs per county (CAFOs per square mile). 
```{r CAFO density}
cbc_table <- as.data.frame(table(swine_df$County.Name))
colnames(cbc_table) <- c("County", "cafo_num")

df <- left_join(df, cbc_table)
df$cafo_num <- ifelse(is.na(df$cafo_num), 0, df$cafo_num)

df$cafo_dens <- df$cafo_num/df$land_area
```

### 7. Population density
```{r popdens}
df$pop_dens <- df$county_pop/df$land_area
```

### 8. Foreign-born
```{r foreign-born}
foreign_df <- foreign_df[116,]
keep_foreign <- vector()
for(i in 1:100){
  keep_foreign[i] <- 6+((i-1)*4)
}
foreign_df <- foreign_df[,keep_foreign]
foreign_df <- pivot_longer(foreign_df, cols=1:100, names_to="County")
colnames(foreign_df) <- c("County", "foreign")
foreign_df$County <- str_remove(foreign_df$County, ".County..North.Carolina..Estimate")
foreign_df$foreign <- str_remove(foreign_df$foreign, ",")
foreign_df$foreign <- as.numeric(foreign_df$foreign)

df <- left_join(df, foreign_df, by="County")
df$foreign <- str_remove(df$foreign, ",")
df$foreign <- as.numeric(df$foreign)

df <- df %>%
  mutate(f_perc=foreign/county_pop*100)

## hard code New Hanover County
df$f_perc[65] <- 6.07
```

### 9. Save
```{r saving dfs}
saveRDS(swine_df, "swine_df.RDS")
saveRDS(df, "df.RDS")
```

