---
title: "EDA and ESDA"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup
```{r load packages}
library(tidyverse)
library(ggplot2)
library(sf)
library(spdep)
library(tmap)
library(viridis)
library(RColorBrewer)
library(corrr)
library(car)
```

```{r load data}
## NC state and county shapefile - 2025, NC OneMap
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp")

df <- readRDS("df.RDS")
swine_df <- readRDS("swine_df.RDS")
```

### 1. EDA

Going to start by looking at all the covariates.
```{r EDA}
summary(df$county_pop); hist(df$county_pop)
summary(df$TB); hist(df$TB)
summary(df$fw); hist(df$fw)
summary(df$ha); hist(df$ha)
summary(df$resp); hist(df$resp)
summary(df$vec); hist(df$vec)
summary(df$beds_incid); hist(df$beds_incid)
summary(df$Percent.in.Poverty); hist(df$Percent.in.Poverty)
summary(df$cafo_dens); hist(df$cafo_dens)

plot(df$Percent.in.Poverty, df$cafo_dens)
```

Do we see correlation among our variables?
```{r correl}
cor.table <- df %>%
  select(-County) %>%
  correlate()
cor.table
```


### 2. Map

For final figures, we would want to do a better job of selecting cutpoints (based more on quantiles) but for now we're just doing five cuts.
```{r make maps}
## pop density
class_pd <- cut(df$pop_dens, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_pd)) +
  scale_fill_brewer(palette="OrRd")

## CAFO density
class_cafo <- cut(df$cafo_dens, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_cafo)) +
  scale_fill_brewer(palette="OrRd")

## poverty
class_pov <- cut(df$Percent.in.Poverty, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_pov)) +
  scale_fill_brewer(palette="OrRd")

## TB
class_TB <- cut(df$TB, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_TB)) +
  scale_fill_brewer(palette="OrRd")

## fw
class_fw <- cut(df$fw, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_fw)) +
  scale_fill_brewer(palette="OrRd")

## ha
class_ha <- cut(df$ha, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_ha)) +
  scale_fill_brewer(palette="OrRd")

## resp
class_resp <- cut(df$resp, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_resp)) +
  scale_fill_brewer(palette="OrRd")

## vec
class_vec <- cut(df$vec, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_vec)) +
  scale_fill_brewer(palette="OrRd")

```

Based on this EDA, it might not be interesting to look at vector-borne diseases or HAIs. 

### 3. Simple linear regression

```{r linreg}
m1 <- lm(TB ~ cafo_dens, data=df)
summary(m1)

m2 <- lm(fw ~ cafo_dens, data=df)
summary(m2)

m3 <- lm(resp ~ cafo_dens, data=df)
summary(m3)
```

```{r multi linreg}
m4 <- lm(TB ~ cafo_dens + pop_dens + beds_incid + Percent.in.Poverty, data=df)
summary(m4)
vif(m4)

hist(resid(m4))
qqPlot(m4) # residuals not normal
plot(resid(m4))

## plot the residuals
df_wresid <- df %>%
  ungroup() %>%
  mutate(olsresid=resid(m4))

class_resids <- cut(df_wresid$olsresid, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_resids)) +
  scale_fill_brewer(palette="OrRd")
```


It looks like TB and CAFO density has a significant relationship. Time to test that more thoroughly.