---
title: "Spatial Analyses"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(sf)
library(spdep)
library(bispdep)
library(spatialreg)
library(splm)
library(plm)
```

```{r load data}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp") 

df_tb <- readRDS("df_tb.RDS")
df_ord <- readRDS("df_ord.RDS")
```

```{r combine data}
NC_tb <- left_join(NC_sf, df_tb, by="County")
NC_ord <- left_join(NC_sf, df_ord, by="County")

NC_tb2017 <- left_join(NC_sf, df_2017tb, by="County")
NC_tb2018 <- left_join(NC_sf, df_2018tb, by="County")
NC_tb2019 <- left_join(NC_sf, df_2019tb, by="County")
NC_tb2020 <- left_join(NC_sf, df_2020tb, by="County")
NC_tb2021 <- left_join(NC_sf, df_2021tb, by="County")
NC_tb2022 <- left_join(NC_sf, df_2022tb, by="County")
NC_tb2023 <- left_join(NC_sf, df_2023tb, by="County")

NC_ord2017 <- left_join(NC_sf, df_2017ord, by="County")
NC_ord2018 <- left_join(NC_sf, df_2018ord, by="County")
NC_ord2019 <- left_join(NC_sf, df_2019ord, by="County")
NC_ord2020 <- left_join(NC_sf, df_2020ord, by="County")
NC_ord2021 <- left_join(NC_sf, df_2021ord, by="County")
NC_ord2022 <- left_join(NC_sf, df_2022ord, by="County")
NC_ord2023 <- left_join(NC_sf, df_2023ord, by="County")
```

We're only going to look at TB and other respiratory illnesses to compare.

The steps we'll take for spatial analyses are:

1.  Defining neighbors and creating spatial weights matrix

2.  Univariate Moran's I - cross-sectionally by year

3.  Bivariate Moran's I - cross-sectionally by year

4.  Prepare panel data

5.  Hausman test

6.  Spatial regression (Durbin) - longitudinal/panel

7.  Visualizing results

### 1. TB

**Defining neighbors and making spatial weights matrix**

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

**Univariate Moran's I**

```{r uni Moran's}
## 2017
### CAFOs
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_tb2017)
plot(lag ~ cafo_psm, NC_tb2017); abline(m1, col="blue"); abline(v=mean(NC_tb2017$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2017$cafo_psm), lty=3, col="grey70")

## get the coefficient
um_cafo <- moran(NC_tb2017$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo

## statistical significance
MC <- moran.mc(NC_tb2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

### FOR TB
## Moran's I scatterplot
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$p_ir)
m2 <- lm(lag ~ p_ir, NC_tb2017)
plot(lag ~ p_ir, NC_tb2017); abline(m2, col="blue"); abline(v=mean(NC_tb2017$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2017$p_ir), lty=3, col="grey70")

## get the coefficient
um_p_ir <- moran(NC_tb2017$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_tb2017$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
### CAFOs
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_tb2020)
plot(lag ~ cafo_psm, NC_tb2020); abline(m1, col="blue"); abline(v=mean(NC_tb2020$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2020$cafo_psm), lty=3, col="grey70")

## get the coefficient
um_cafo <- moran(NC_tb2020$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo

## statistical significance
MC <- moran.mc(NC_tb2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

### FOR TB
## Moran's I scatterplot
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$p_ir)
m2 <- lm(lag ~ p_ir, NC_tb2020)
plot(lag ~ p_ir, NC_tb2020); abline(m2, col="blue"); abline(v=mean(NC_tb2020$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2020$p_ir), lty=3, col="grey70")

## get the coefficient
um_p_ir <- moran(NC_tb2020$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_tb2020$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
### CAFOs
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_tb2023)
plot(lag ~ cafo_psm, NC_tb2023); abline(m1, col="blue"); abline(v=mean(NC_tb2023$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2023$cafo_psm), lty=3, col="grey70")

## get the coefficient
um_cafo <- moran(NC_tb2023$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo

## statistical significance
MC <- moran.mc(NC_tb2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

### FOR TB
## Moran's I scatterplot
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$p_ir)
m2 <- lm(lag ~ p_ir, NC_tb2023)
plot(lag ~ p_ir, NC_tb2023); abline(m2, col="blue"); abline(v=mean(NC_tb2023$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2023$p_ir), lty=3, col="grey70")

## get the coefficient
um_p_ir <- moran(NC_tb2023$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_tb2023$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

The CAFOs across the three years exhibit higher I than TB, which is below 0.2 for all three years but varies across them. 2020 has the lowest clustering of TB.

**Bivariate Moran's I**

```{r bi Moran's}
## 2017
## get the coefficient
bm_TB2017 <- moran.bi(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights)
bm_TB2017

## statistical significance
MC <- moranbi.mc(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
## get the coefficient
bm_TB2020 <- moran.bi(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights)
bm_TB2020

## statistical significance
MC <- moranbi.mc(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
## get the coefficient
bm_TB2023 <- moran.bi(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights)
bm_TB2023

## statistical significance
MC <- moranbi.mc(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Like with the individual Moran's I, 2020 has the lowest I.

**Prepare panel data**

```{r prepare panel data}
df_tb <- df_tb %>%
  select(County, year, type, cat_n, County.Pop.Adjusted, p_ir, count, cafo_psm, perc_unins, perc_pov, n_foreign, Total.Land.Area, pop_dens, perc_foreign) %>%
  arrange(County, year)

panel_tb <- pdata.frame(df_tb, index = c("County", "year"))
```

Since we see spatial autocorrelation, we do a spatial regression

```{r spatial regression}
## OLS
o1 <- lm(
  TB ~ cafo_psm + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_tb
)
summary(o1)
saveRDS(o1, "o1.RDS")

## SAR
m3 <- lagsarlm(
  TB ~ cafo_psm + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_tb,
  listw = weights
)
summary(m3)
saveRDS(m3, "m3.RDS")

# get estimates of direct, indirect, and total effect of each variable because cannot be compared directly to OLS coefficients
m3_impacts <- impacts(m3, listw=weights, R=999)
m3_impacts
summary(m3_impacts, zstats=TRUE)

## SER
m4 <- errorsarlm(
  TB ~ cafo_psm + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df, 
  listw = weights
)
summary(m4) # CAN be directly compared to OLS coefficients
saveRDS(m4, "m4.RDS")

## compare
AIC(m3); AIC(m4)
```

Some more information on how to interpret the results:

-   For the SAR model, the effect of a change in covariates spatially propagates through neighbors, neighbors of neighbors, etc. So, we can't interpret the raw $\beta$ values. So, we use the `impacts` function to decompose the direct and indirect impacts (marginal effects). **Direct impacts** are the average effect of changing a covariate within its own unit; **indirect impacts** are for the impact of the change on other units. The **total impacts** are direct + indirect. Example: if the direct impact of income on house prices is 0.5 and the indirect is 0.2, then a 1-unit increase in income in a neighborhood increases its own house price by 0.5 and increases that of nearby homes by 0.2. The total effect is 0.7.
-   For the SER model, the $\beta$ coefficients are interpreted just like in OLS.

CAFO density is highly significant even after accounting for spatial autocorrelation, population density, % poverty, and hospital beds per person. Model 1 (SAR) is the better model.

Now we compare to what that looks like with other respiratory illnesses.

### 2. Other respiratory illnesses

Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

Univariate Moran's I

```{r uni Moran's}
## Moran's I scatterplot
NC_sf$lag <- lag.listw(weights, NC_sf$resp)
m5 <- lm(lag ~ resp, NC_sf)
plot(lag ~ resp, NC_sf)
abline(m5, col="blue")
abline(v=mean(NC_sf$resp), lty=3, col="grey70")
abline(h=mean(NC_sf$resp), lty=3, col="grey70")

## get the coefficient
um_resp <- moran(NC_sf$resp, listw=weights, n=length(nb), S0=Szero(weights))
um_resp

## statistical significance
MC <- moran.mc(NC_sf$resp, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Bivariate Moran's I

```{r bi Moran's}
## get the coefficient
bm_resp <- moran.bi(NC_sf$resp, NC_sf$cafo_dens, weights)
bm_resp

## statistical significance
MC <- moranbi.mc(NC_sf$resp, NC_sf$cafo_dens, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Since we see spatial autocorrelation, we do a spatial regression

```{r spatial regression}
## OLS
o2 <- lm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df
)
summary(o2)
saveRDS(o2, "o2.RDS")

## SAR
m6 <- lagsarlm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_sf,
  listw = weights
)
summary(m6)
saveRDS(m6, "m6.RDS")

# get estimates of direct, indirect, and total effect of each variable because cannot be compared directly to OLS coefficients
m6_impacts <- impacts(m6, listw=weights, R=999)
m6_impacts
summary(m6_impacts, zstats=TRUE)

## SER
m7 <- errorsarlm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df, 
  listw = weights
)
summary(m7) # CAN be directly compared to OLS coefficients
saveRDS(m7, "m7.RDS")

## compare
AIC(m6); AIC(m7)
```
