---
title: "Spatial Analyses"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(spdep)
library(bispdep)
library(spatialreg)
```

```{r load data}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp") 

df <- readRDS("df.RDS")
swine_df <- readRDS("swine_df.RDS")
```

```{r combine data}
NC_all <- left_join(NC_sf, df, by="County")
```

We're only going to look at TB and other respiratory illnesses to compare.

The steps we'll take for spatial analyses are:

1.  Defining neighbors and creating spatial weights matrix

2.  Univariate Moran's I

3.  Bivariate Moran's I

4.  Spatial regression

5.  Visualizing results

6.  Mediation analyses

### 1. TB

Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_all, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

Univariate Moran's I

```{r uni Moran's}
### FOR CAFOs
NC_all$lag <- lag.listw(weights, NC_all$cafo_dens)
m1 <- lm(lag ~ cafo_dens, NC_all)
plot(lag ~ cafo_dens, NC_all)
abline(m1, col="blue")
abline(v=mean(NC_all$cafo_dens), lty=3, col="grey70")
abline(h=mean(NC_all$cafo_dens), lty=3, col="grey70")

## get the coefficient
um_cafo <- moran(NC_all$cafo_dens, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo

## statistical significance
MC <- moran.mc(NC_all$cafo_dens, weights, nsim=999)
MC
plot(MC)
MC$p.value

### FOR TB
## Moran's I scatterplot
NC_all$lag <- lag.listw(weights, NC_all$TB)
m2 <- lm(lag ~ TB, NC_all)
plot(lag ~ TB, NC_all)
abline(m2, col="blue")
abline(v=mean(NC_all$TB), lty=3, col="grey70")
abline(h=mean(NC_all$TB), lty=3, col="grey70")

## get the coefficient
um_TB <- moran(NC_all$TB, listw=weights, n=length(nb), S0=Szero(weights))
um_TB

## statistical significance
MC <- moran.mc(NC_all$TB, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Bivariate Moran's I

```{r bi Moran's}
## get the coefficient
bm_TB <- moran.bi(NC_all$TB, NC_all$cafo_dens, weights)
bm_TB

## statistical significance
MC <- moranbi.mc(NC_all$TB, NC_all$cafo_dens, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Since we see spatial autocorrelation, we do a spatial regression

```{r spatial regression}
## OLS
o1 <- lm(
  TB ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_all
)
summary(o1)
saveRDS(o1, "o1.RDS")

## SAR
m3 <- lagsarlm(
  TB ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_all,
  listw = weights
)
summary(m3)
saveRDS(m3, "m3.RDS")

# get estimates of direct, indirect, and total effect of each variable because cannot be compared directly to OLS coefficients
m3_impacts <- impacts(m3, listw=weights, R=999)
m3_impacts
summary(m3_impacts, zstats=TRUE)

## SER
m4 <- errorsarlm(
  TB ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df, 
  listw = weights
)
summary(m4) # CAN be directly compared to OLS coefficients
saveRDS(m4, "m4.RDS")

## compare
AIC(m3); AIC(m4)
```

Some more information on how to interpret the results:

-   For the SAR model, the effect of a change in covariates spatially propagates through neighbors, neighbors of neighbors, etc. So, we can't interpret the raw $\beta$ values. So, we use the `impacts` function to decompose the direct and indirect impacts (marginal effects). **Direct impacts** are the average effect of changing a covariate within its own unit; **indirect impacts** are for the impact of the change on other units. The **total impacts** are direct + indirect. Example: if the direct impact of income on house prices is 0.5 and the indirect is 0.2, then a 1-unit increase in income in a neighborhood increases its own house price by 0.5 and increases that of nearby homes by 0.2. The total effect is 0.7.
-   For the SER model, the $\beta$ coefficients are interpreted just like in OLS.

CAFO density is highly significant even after accounting for spatial autocorrelation, population density, % poverty, and hospital beds per person. Model 1 (SAR) is the better model.

Now we compare to what that looks like with other respiratory illnesses.

### 2. Other respiratory illnesses

Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_all, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

Univariate Moran's I

```{r uni Moran's}
## Moran's I scatterplot
NC_all$lag <- lag.listw(weights, NC_all$resp)
m5 <- lm(lag ~ resp, NC_all)
plot(lag ~ resp, NC_all)
abline(m5, col="blue")
abline(v=mean(NC_all$resp), lty=3, col="grey70")
abline(h=mean(NC_all$resp), lty=3, col="grey70")

## get the coefficient
um_resp <- moran(NC_all$resp, listw=weights, n=length(nb), S0=Szero(weights))
um_resp

## statistical significance
MC <- moran.mc(NC_all$resp, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Bivariate Moran's I

```{r bi Moran's}
## get the coefficient
bm_resp <- moran.bi(NC_all$resp, NC_all$cafo_dens, weights)
bm_resp

## statistical significance
MC <- moranbi.mc(NC_all$resp, NC_all$cafo_dens, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Since we see spatial autocorrelation, we do a spatial regression

```{r spatial regression}
## OLS
o2 <- lm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df
)
summary(o2)
saveRDS(o2, "o2.RDS")

## SAR
m6 <- lagsarlm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = NC_all,
  listw = weights
)
summary(m6)
saveRDS(m6, "m6.RDS")

# get estimates of direct, indirect, and total effect of each variable because cannot be compared directly to OLS coefficients
m6_impacts <- impacts(m6, listw=weights, R=999)
m6_impacts
summary(m6_impacts, zstats=TRUE)

## SER
m7 <- errorsarlm(
  resp ~ cafo_dens + pop_dens + Percent.in.Poverty + beds_pc + f_perc,
  data = df, 
  listw = weights
)
summary(m7) # CAN be directly compared to OLS coefficients
saveRDS(m7, "m7.RDS")

## compare
AIC(m6); AIC(m7)
```
