---
title: "Spatial Analyses"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(car)
library(sf)
library(spdep)
library(bispdep)
library(spatialreg)
library(splm)
library(plm)
library(GWmodel)
```

```{r load data}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp") 

df_tb <- readRDS("df_tb.RDS")
df_ord <- readRDS("df_ord.RDS")
```

```{r combine data}
NC_tb <- left_join(NC_sf, df_tb, by="County")
NC_ord <- left_join(NC_sf, df_ord, by="County")

NC_tb2017 <- left_join(NC_sf, df_2017tb, by="County")
NC_tb2018 <- left_join(NC_sf, df_2018tb, by="County")
NC_tb2019 <- left_join(NC_sf, df_2019tb, by="County")
NC_tb2020 <- left_join(NC_sf, df_2020tb, by="County")
NC_tb2021 <- left_join(NC_sf, df_2021tb, by="County")
NC_tb2022 <- left_join(NC_sf, df_2022tb, by="County")
NC_tb2023 <- left_join(NC_sf, df_2023tb, by="County")

NC_ord2017 <- left_join(NC_sf, df_2017ord, by="County")
NC_ord2018 <- left_join(NC_sf, df_2018ord, by="County")
NC_ord2019 <- left_join(NC_sf, df_2019ord, by="County")
NC_ord2020 <- left_join(NC_sf, df_2020ord, by="County")
NC_ord2021 <- left_join(NC_sf, df_2021ord, by="County")
NC_ord2022 <- left_join(NC_sf, df_2022ord, by="County")
NC_ord2023 <- left_join(NC_sf, df_2023ord, by="County")
```

We're only going to look at TB and other respiratory illnesses to compare.

The steps we'll take for spatial analyses are:

1.  Defining neighbors and creating spatial weights matrix

2.  Univariate global Moran's I - cross-sectionally by year

3.  Bivariate Moran's I - cross-sectionally by year

4.  Local Moran's I - cross-sectionally by year

5.  Prepare panel data

6.  Spatial regression - SEM

7.  Troubleshooting

8.  GTWR

9.  Visualizing results



S
### 1. TB

#### Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

**Univariate Moran's I**

We will do these for TB (outcome) and CAFOs (key predictor). Start by getting the lags and residuals to plot and check for constant variance. Then get the global Moran's I and its p-value. Then get the Moran's I for the residuals. Lastly, calculate local Moran's.

```{r year 2017}
### CAFOs ----
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_tb2017)
plot(lag ~ cafo_psm, NC_tb2017); abline(m1, col="blue"); abline(v=mean(NC_tb2017$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2017$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo17 <- tibble(
  res = m1$residuals,
  fitted = m1$fitted.values
)

ggplot(data = r_cafo17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo17 <- moran(NC_tb2017$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo17

## statistical significance
MC_cafo17 <- moran.mc(NC_tb2017$cafo_psm, weights, nsim=999)
MC_cafo17
plot(MC_cafo17)
MC_cafo17$p.value

## Moran's on residuals
rm_cafo17 <- lm.morantest(m1, listw = weights, alternative = "greater")
rm_cafo17

### TB ----
## Moran's I scatterplot
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$p_ir)
m2 <- lm(lag ~ p_ir, NC_tb2017)
plot(lag ~ p_ir, NC_tb2017); abline(m2, col="blue"); abline(v=mean(NC_tb2017$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2017$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb17 <- tibble(
  res = m2$residuals,
  fitted = m2$fitted.values
)

ggplot(data = r_tb17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_p_ir <- moran(NC_tb2017$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_tb2017$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## Moran's on residuals
rm_tb17 <- lm.morantest(m2, listw = weights, alternative = "greater")
rm_tb17

## Lagrange tests for TB
lm.LMtests(m2, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m2, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2020}
### CAFOs ----
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$cafo_psm)
m3 <- lm(lag ~ cafo_psm, NC_tb2020)
plot(lag ~ cafo_psm, NC_tb2020); abline(m3, col="blue"); abline(v=mean(NC_tb2020$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2020$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo20 <- tibble(
  res = m3$residuals,
  fitted = m3$fitted.values
)

ggplot(data = r_cafo20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo20 <- moran(NC_tb2020$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo20

## statistical significance
MC_cafo20 <- moran.mc(NC_tb2020$cafo_psm, weights, nsim=999)
MC_cafo20
plot(MC_cafo20)
MC_cafo20$p.value

## Moran's on residuals
rm_cafo20 <- lm.morantest(m3, listw = weights, alternative = "greater")
rm_cafo20

### TB ----
## Moran's I scatterplot
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$p_ir)
m4 <- lm(lag ~ p_ir, NC_tb2020)
plot(lag ~ p_ir, NC_tb2020); abline(m4, col="blue"); abline(v=mean(NC_tb2020$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2020$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb20 <- tibble(
  res = m4$residuals,
  fitted = m4$fitted.values
)

ggplot(data = r_tb20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_tb20 <- moran(NC_tb2020$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_tb20

## statistical significance
MC_tb20 <- moran.mc(NC_tb2020$p_ir, weights, nsim=999)
MC_tb20
plot(MC_tb20)
MC_tb20$p.value

## Moran's on residuals
rm_tb20 <- lm.morantest(m4, listw = weights, alternative = "greater")
rm_tb20

## Lagrange tests for TB
lm.LMtests(m4, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m4, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2023}
### CAFOs ----
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$cafo_psm)
m5 <- lm(lag ~ cafo_psm, NC_tb2023)
plot(lag ~ cafo_psm, NC_tb2023); abline(m5, col="blue"); abline(v=mean(NC_tb2023$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2023$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo23 <- tibble(
  res = m5$residuals,
  fitted = m5$fitted.values
)

ggplot(data = r_cafo23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo23 <- moran(NC_tb2023$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo23

## statistical significance
MC_cafo23 <- moran.mc(NC_tb2023$cafo_psm, weights, nsim=999)
MC_cafo23
plot(MC_cafo23)
MC_cafo23$p.value

## Moran's on residuals
rm_cafo23 <- lm.morantest(m5, listw = weights, alternative = "greater")
rm_cafo23

### TB ----
## Moran's I scatterplot
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$p_ir)
m6 <- lm(lag ~ p_ir, NC_tb2023)
plot(lag ~ p_ir, NC_tb2023); abline(m6, col="blue"); abline(v=mean(NC_tb2023$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2023$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb23 <- tibble(
  res = m6$residuals,
  fitted = m6$fitted.values
)

ggplot(data = r_tb23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_tb23 <- moran(NC_tb2023$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_tb23

## statistical significance
MC_tb23 <- moran.mc(NC_tb2023$p_ir, weights, nsim=999)
MC_tb23
plot(MC_tb23)
MC_tb23$p.value

## Moran's on residuals
rm_tb23 <- lm.morantest(m6, listw = weights, alternative = "greater")
rm_tb23

## Lagrange tests for TB
lm.LMtests(m6, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m6, weights, test = c("RLMerr", "RLMlag"))
```

The CAFOs across the three years exhibit higher I than TB, which is below 0.2 for all three years but varies across them. 2020 has the lowest clustering of TB.

There's also a pretty distinct clustering of residuals and lack of constant variance with CAFOs and TB.

#### Bivariate Moran's I

```{r bi Moran's}
## 2017
## get the coefficient
bm_TB2017 <- moran.bi(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights)
bm_TB2017

## statistical significance
MC <- moranbi.mc(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
## get the coefficient
bm_TB2020 <- moran.bi(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights)
bm_TB2020

## statistical significance
MC <- moranbi.mc(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
## get the coefficient
bm_TB2023 <- moran.bi(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights)
bm_TB2023

## statistical significance
MC <- moranbi.mc(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Like with the individual Moran's I, 2020 has the lowest I.

#### Local Moran's I

```{r local}
## 2017 ----
local_m2 <- localmoran(df_2017tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2017 <- order(df_2017tb$County)
local_m2
printCoefmat(data.frame(local_m2[county_2017,], row.names=df_2017tb$County[county_2017]), check.names=FALSE)

localg_m2 <- localG(df_2017tb$p_ir, listw = weights)
localg_m2
NC_tb2017$GetisOrd <- localg_m2
NC_tb2017 <- NC_tb2017 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2017) +
  geom_sf(aes(fill = Gi_cat))

## 2020 ----
local_m4 <- localmoran(df_2020tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2020 <- order(df_2020tb$County)
local_m4
printCoefmat(data.frame(local_m4[county_2020,], row.names=df_2020tb$County[county_2020]), check.names=FALSE)

localg_m4 <- localG(df_2020tb$p_ir, listw = weights)
localg_m4
NC_tb2020$GetisOrd <- localg_m4
NC_tb2020 <- NC_tb2020 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2020) +
  geom_sf(aes(fill = Gi_cat))

## 2023 ----
local_m6 <- localmoran(df_2023tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2023 <- order(df_2023tb$County)
local_m6
printCoefmat(data.frame(local_m6[county_2023,], row.names=df_2023tb$County[county_2023]), check.names=FALSE)

localg_m6 <- localG(df_2023tb$p_ir, listw = weights)
localg_m6
NC_tb2023$GetisOrd <- localg_m6
NC_tb2023 <- NC_tb2023 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2023) +
  geom_sf(aes(fill = Gi_cat))
```

#### Prepare panel data

```{r prepare panel data}
df_tb <- df_tb %>%
  select(County, year, type, cat_n, County.Pop.Adjusted, p_ir, count, cafo_psm, perc_unins, perc_pov, n_foreign, Total.Land.Area, pop_dens, perc_foreign) %>%
  arrange(County, year)

panel_tb <- pdata.frame(df_tb, index = c("County", "year"))
```

#### Spatial regression

We're doing a spatial Durbin model, which accounts for autocorrelated outcomes and residuals.

```{r SEM}
sm1 <- spml(
  p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = panel_tb,
  listw = weights,
  model = "within",      # fixed effect
  effect = "twoway",     # both time and unit heterogeneity
  lag = TRUE,
  Durbin = TRUE
)
summary(sm1)
splm::impacts(sm1, listw = weights, time = 7, R = 500)

sm2 <- spml(
  p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = panel_tb,
  listw = weights,
  model = "random",      # random effect
  lag = TRUE,
  Durbin = TRUE
)
summary(sm2)
```

We use Hausman tests to check whether the coefficients in the fixed vs random effect models differ significantly. If significant, use the FE model, because it captures group-level bias and omitted variables better, and assumes correlation between unobserved and observed characteristics

```{r Hausman}
sphtest(sm1, sm2)
```

The Hausman test tells us that the fixed effects model is the right way to go. The significance we see in the random effects model is likely due to perceived between-county associations which are actually just spatially clustered variables.

Nothing being significant is a little weird, though. It's possible that there is too little within-county, per-year variation in some of our variables (aka we have unexpectedly time-invariant predictors). Let's check that.

#### Troubleshooting

Let's compare within-county and between-county variation. Trying coefficient of variation to look at within-county variation (aka variation over time).

```{r eyeball}
summary(panel_tb)

## TB
ts_tb <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(p_ir),
            mean = mean(p_ir),
            cv = within_sd/mean)
ts_tb$cv <- ifelse(is.na(ts_tb$cv), 0, ts_tb$cv)
hist(ts_tb$cv)

## CAFO count
ts_cc <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(count),
            mean = mean(count),
            cv = within_sd/mean)
ts_cc$cv <- ifelse(is.na(ts_cc$cv), 0, ts_cc$cv)
hist(ts_cc$cv)

## CAFO density
ts_cafo <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(cafo_psm),
            mean = mean(cafo_psm),
            cv = within_sd/mean)
ts_cafo$cv <- ifelse(is.na(ts_cafo$cv), 0, ts_cafo$cv)
hist(ts_cafo$cv)

## population density
ts_pd <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(pop_dens),
            mean = mean(pop_dens),
            cv = within_sd/mean)
ts_pd$cv <- ifelse(is.na(ts_pd$cv), 0, ts_pd$cv)
hist(ts_pd$cv)

## percent poverty
ts_pov <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(perc_pov),
            mean = mean(perc_pov),
            cv = within_sd/mean)
ts_pov$cv <- ifelse(is.na(ts_pov$cv), 0, ts_pov$cv)
hist(ts_pov$cv)

## percent poverty
ts_unins <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(perc_unins),
            mean = mean(perc_unins),
            cv = within_sd/mean)
ts_unins$cv <- ifelse(is.na(ts_unins$cv), 0, ts_unins$cv)
hist(ts_unins$cv)
```

It looks like there's low within-county variation/change over time for most of the variables, so there really isn't much variation for us to model.

Let's try a different approach.

#### GTWR

Geographically weighted models (GWRs) differ from the other spatial regressions because it allows for nonstationarity, where regression coefficients can vary spatially. Aka, the relationship between the coefficients and TB may differ by county, and it produces unique coefficients for each county instead of one global set of coefficients.

Let's see how this goes.

First, the data must be an sf object. We already have an sf version of the data; let's get it in a format they want. Specifically, we projected coordinates and county centroids.

```{r data prep}
gtwr_tb <- st_transform(NC_tb, crs = 3857)
tb_coords <- st_coordinates(st_centroid(gtwr_tb))
```

Bandwidth selection, aka choosing the size of the "neighborhood" of each location in the regression. Looking for the lowest AIC.

```{r bandwidth selection}
bw_tb <- bw.gtwr(
  formula = p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = gtwr_tb,
  obs.tv = gtwr_tb$year,
  approach = "AICc",
  kernel = "bisquare",
  adaptive = TRUE
)
```

Fit model.

```{r GTWR model fitting}
gm_tb <- gtwr(
  formula = p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = gtwr_tb,
  obs.tv = gtwr_tb$year,
  st.bw = bw_tb,
  kernel = "bisquare",
  adaptive = TRUE
)
```

Results time.

```{r GTWR results}
gm_tb # will give you global results and summaries of each of the predictors

local_coefs <- gm_tb$SDF
local_coefs$County <- NC_tb$County
summary(local_coefs$cafo_psm)
```

If you want to check the significance of the nonstationarity, compare the IQR of the parameter estimates with 2 times the SE of the OLS estimates (Fotheringham, Brunsdon, and Charlton 2002; Fotheringham, Crespo, and Yao 2015); larger IQRs indicate significant spatial nonstationarity, and that global/average descriptions of relationships between variables is insufficient.

```{r GTWR significance}
summary(gm_tb$lm)

IQR(local_coefs$cafo_psm); 2*0.9793252     # IQR > 2*OLS SE
IQR(local_coefs$pop_dens); 2*0.0002714     # IQR < 2*OLS SE
IQR(local_coefs$perc_unins); 2*0.0364409   # IQR > 2*OLS SE
IQR(local_coefs$perc_pov); 2*0.0190205     # IQR > 2*OLS SE
```

Looks like there's quite a bit of non-stationarity in the four predictors we have.

#### Visualizing results

Local Moran's I and LISA clusters





### 2. ORDs

#### Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

**Univariate Moran's I**

We will do these for ORD (outcome) and CAFOs (key predictor). Start by getting the lags and residuals to plot and check for constant variance. Then get the global Moran's I and its p-value. Then get the Moran's I for the residuals. Lastly, calculate local Moran's.

```{r year 2017}
### CAFOs ----
NC_ord2017$lag <- lag.listw(weights, NC_ord2017$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_ord2017)
plot(lag ~ cafo_psm, NC_ord2017); abline(m1, col="blue"); abline(v=mean(NC_ord2017$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2017$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo17 <- tibble(
  res = m1$residuals,
  fitted = m1$fitted.values
)

ggplot(data = r_cafo17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo17 <- moran(NC_ord2017$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo17

## statistical significance
MC_cafo17 <- moran.mc(NC_ord2017$cafo_psm, weights, nsim=999)
MC_cafo17
plot(MC_cafo17)
MC_cafo17$p.value

## Moran's on residuals
rm_cafo17 <- lm.morantest(m1, listw = weights, alternative = "greater")
rm_cafo17

### ORD ----
## Moran's I scatterplot
NC_ord2017$lag <- lag.listw(weights, NC_ord2017$p_ir)
m2 <- lm(lag ~ p_ir, NC_ord2017)
plot(lag ~ p_ir, NC_ord2017); abline(m2, col="blue"); abline(v=mean(NC_ord2017$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2017$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord17 <- tibble(
  res = m2$residuals,
  fitted = m2$fitted.values
)

ggplot(data = r_ord17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_p_ir <- moran(NC_ord2017$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_ord2017$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## Moran's on residuals
rm_ord17 <- lm.morantest(m2, listw = weights, alternative = "greater")
rm_ord17

## Lagrange tests for ORD
lm.LMtests(m2, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m2, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2020}
### CAFOs ----
NC_ord2020$lag <- lag.listw(weights, NC_ord2020$cafo_psm)
m3 <- lm(lag ~ cafo_psm, NC_ord2020)
plot(lag ~ cafo_psm, NC_ord2020); abline(m3, col="blue"); abline(v=mean(NC_ord2020$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2020$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo20 <- tibble(
  res = m3$residuals,
  fitted = m3$fitted.values
)

ggplot(data = r_cafo20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo20 <- moran(NC_ord2020$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo20

## statistical significance
MC_cafo20 <- moran.mc(NC_ord2020$cafo_psm, weights, nsim=999)
MC_cafo20
plot(MC_cafo20)
MC_cafo20$p.value

## Moran's on residuals
rm_cafo20 <- lm.morantest(m3, listw = weights, alternative = "greater")
rm_cafo20

### ORD ----
## Moran's I scatterplot
NC_ord2020$lag <- lag.listw(weights, NC_ord2020$p_ir)
m4 <- lm(lag ~ p_ir, NC_ord2020)
plot(lag ~ p_ir, NC_ord2020); abline(m4, col="blue"); abline(v=mean(NC_ord2020$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2020$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord20 <- tibble(
  res = m4$residuals,
  fitted = m4$fitted.values
)

ggplot(data = r_ord20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_ord20 <- moran(NC_ord2020$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_ord20

## statistical significance
MC_ord20 <- moran.mc(NC_ord2020$p_ir, weights, nsim=999)
MC_ord20
plot(MC_ord20)
MC_ord20$p.value

## Moran's on residuals
rm_ord20 <- lm.morantest(m4, listw = weights, alternative = "greater")
rm_ord20

## Lagrange tests for ORD
lm.LMtests(m4, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m4, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2023}
### CAFOs ----
NC_ord2023$lag <- lag.listw(weights, NC_ord2023$cafo_psm)
m5 <- lm(lag ~ cafo_psm, NC_ord2023)
plot(lag ~ cafo_psm, NC_ord2023); abline(m5, col="blue"); abline(v=mean(NC_ord2023$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2023$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo23 <- tibble(
  res = m5$residuals,
  fitted = m5$fitted.values
)

ggplot(data = r_cafo23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo23 <- moran(NC_ord2023$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo23

## statistical significance
MC_cafo23 <- moran.mc(NC_ord2023$cafo_psm, weights, nsim=999)
MC_cafo23
plot(MC_cafo23)
MC_cafo23$p.value

## Moran's on residuals
rm_cafo23 <- lm.morantest(m5, listw = weights, alternative = "greater")
rm_cafo23

### ORD ----
## Moran's I scatterplot
NC_ord2023$lag <- lag.listw(weights, NC_ord2023$p_ir)
m6 <- lm(lag ~ p_ir, NC_ord2023)
plot(lag ~ p_ir, NC_ord2023); abline(m6, col="blue"); abline(v=mean(NC_ord2023$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2023$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord23 <- tibble(
  res = m6$residuals,
  fitted = m6$fitted.values
)

ggplot(data = r_ord23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_ord23 <- moran(NC_ord2023$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_ord23

## statistical significance
MC_ord23 <- moran.mc(NC_ord2023$p_ir, weights, nsim=999)
MC_ord23
plot(MC_ord23)
MC_ord23$p.value

## Moran's on residuals
rm_ord23 <- lm.morantest(m6, listw = weights, alternative = "greater")
rm_ord23

## Lagrange tests for ORD
lm.LMtests(m6, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m6, weights, test = c("RLMerr", "RLMlag"))
```

We see apparent clustering as well.

#### Bivariate Moran's I

```{r bi Moran's}
## 2017
## get the coefficient
bm_ORD2017 <- moran.bi(NC_ord2017$p_ir, NC_ord2017$cafo_psm, weights)
bm_ORD2017

## statistical significance
MC <- moranbi.mc(NC_ord2017$p_ir, NC_ord2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
## get the coefficient
bm_ORD2020 <- moran.bi(NC_ord2020$p_ir, NC_ord2020$cafo_psm, weights)
bm_ORD2020

## statistical significance
MC <- moranbi.mc(NC_ord2020$p_ir, NC_ord2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
## get the coefficient
bm_ORD2023 <- moran.bi(NC_ord2023$p_ir, NC_ord2023$cafo_psm, weights)
bm_ORD2023

## statistical significance
MC <- moranbi.mc(NC_ord2023$p_ir, NC_ord2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value
```

Very little if any bivariate autocorrelation.

#### Local Moran's I

```{r local}
## 2017 ----
local_m2 <- localmoran(df_2017ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2017 <- order(df_2017ord$County)
local_m2
printCoefmat(data.frame(local_m2[county_2017,], row.names=df_2017ord$County[county_2017]), check.names=FALSE)

localg_m2 <- localG(df_2017ord$p_ir, listw = weights)
localg_m2
NC_ord2017$GetisOrd <- localg_m2
NC_ord2017 <- NC_ord2017 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2017) +
  geom_sf(aes(fill = Gi_cat))

## 2020 ----
local_m4 <- localmoran(df_2020ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2020 <- order(df_2020ord$County)
local_m4
printCoefmat(data.frame(local_m4[county_2020,], row.names=df_2020ord$County[county_2020]), check.names=FALSE)

localg_m4 <- localG(df_2020ord$p_ir, listw = weights)
localg_m4
NC_ord2020$GetisOrd <- localg_m4
NC_ord2020 <- NC_ord2020 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2020) +
  geom_sf(aes(fill = Gi_cat))

## 2023 ----
local_m6 <- localmoran(df_2023ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2023 <- order(df_2023ord$County)
local_m6
printCoefmat(data.frame(local_m6[county_2023,], row.names=df_2023ord$County[county_2023]), check.names=FALSE)

localg_m6 <- localG(df_2023ord$p_ir, listw = weights)
localg_m6
NC_ord2023$GetisOrd <- localg_m6
NC_ord2023 <- NC_ord2023 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2023) +
  geom_sf(aes(fill = Gi_cat))
```
