---
title: "Spatial Analyses"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup

```{r load packages}
library(tidyverse)
library(ggplot2)
library(car)
library(sf)
library(spdep)
library(bispdep)
library(spatialreg)
library(splm)
library(plm)
```

```{r load data}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp") 

df_tb <- readRDS("df_tb.RDS")
df_ord <- readRDS("df_ord.RDS")
```

```{r combine data}
NC_tb <- left_join(NC_sf, df_tb, by="County")
NC_ord <- left_join(NC_sf, df_ord, by="County")

NC_tb2017 <- left_join(NC_sf, df_2017tb, by="County")
NC_tb2018 <- left_join(NC_sf, df_2018tb, by="County")
NC_tb2019 <- left_join(NC_sf, df_2019tb, by="County")
NC_tb2020 <- left_join(NC_sf, df_2020tb, by="County")
NC_tb2021 <- left_join(NC_sf, df_2021tb, by="County")
NC_tb2022 <- left_join(NC_sf, df_2022tb, by="County")
NC_tb2023 <- left_join(NC_sf, df_2023tb, by="County")

NC_ord2017 <- left_join(NC_sf, df_2017ord, by="County")
NC_ord2018 <- left_join(NC_sf, df_2018ord, by="County")
NC_ord2019 <- left_join(NC_sf, df_2019ord, by="County")
NC_ord2020 <- left_join(NC_sf, df_2020ord, by="County")
NC_ord2021 <- left_join(NC_sf, df_2021ord, by="County")
NC_ord2022 <- left_join(NC_sf, df_2022ord, by="County")
NC_ord2023 <- left_join(NC_sf, df_2023ord, by="County")
```

We're only going to look at TB and other respiratory illnesses to compare.

The steps we'll take for spatial analyses are:

1.  Defining neighbors and creating spatial weights matrix

2.  Univariate global Moran's I - cross-sectionally by year

3.  Bivariate Moran's I and Lee's L - cross-sectionally by year

4.  Local Moran's I - cross-sectionally by year

5.  Prepare panel data

6.  Spatial regression - SEM

7.  Troubleshooting

8.  SEM - pooled

9.  Visualizing results - see `tables and figures.RMD` for final visualizations

### 1. TB

#### Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

#### Univariate Moran's I

We will do these for TB (outcome) and CAFOs (key predictor). Start by getting the lags and residuals to plot and check for constant variance. Then get the global Moran's I and its p-value. Then get the Moran's I for the residuals of the lag OLSs, calculate Lagrange, and compute residuals for the full model for Moran's I on those residuals.

```{r year 2017}
### CAFOs ----
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_tb2017)
plot(lag ~ cafo_psm, NC_tb2017); abline(m1, col="blue"); abline(v=mean(NC_tb2017$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2017$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo17 <- tibble(
  res = m1$residuals,
  fitted = m1$fitted.values
)

ggplot(data = r_cafo17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo17 <- moran(NC_tb2017$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo17

## statistical significance
MC_cafo17 <- moran.mc(NC_tb2017$cafo_psm, weights, nsim=999)
MC_cafo17
plot(MC_cafo17)
MC_cafo17$p.value

## Moran's on residuals
rm_cafo17 <- lm.morantest(m1, listw = weights, alternative = "greater")
rm_cafo17

### TB ----
## Moran's I scatterplot
NC_tb2017$lag <- lag.listw(weights, NC_tb2017$p_ir)
m2 <- lm(lag ~ p_ir, NC_tb2017)
plot(lag ~ p_ir, NC_tb2017); abline(m2, col="blue"); abline(v=mean(NC_tb2017$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2017$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb17 <- tibble(
  res = m2$residuals,
  fitted = m2$fitted.values
)

ggplot(data = r_tb17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_p_ir <- moran(NC_tb2017$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_tb2017$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## Moran's on residuals
rm_tb17 <- lm.morantest(m2, listw = weights, alternative = "greater")
rm_tb17

## Lagrange tests for TB
lm.LMtests(m2, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m2, weights, test = c("RLMerr", "RLMlag"))

## full residual Moran's
m2017 <- lm(p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins, data = NC_tb2017)
summary(m2017)
lm.morantest(m2017, listw = weights, alternative = "greater")
```

```{r year 2020}
### CAFOs ----
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$cafo_psm)
m3 <- lm(lag ~ cafo_psm, NC_tb2020)
plot(lag ~ cafo_psm, NC_tb2020); abline(m3, col="blue"); abline(v=mean(NC_tb2020$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2020$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo20 <- tibble(
  res = m3$residuals,
  fitted = m3$fitted.values
)

ggplot(data = r_cafo20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo20 <- moran(NC_tb2020$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo20

## statistical significance
MC_cafo20 <- moran.mc(NC_tb2020$cafo_psm, weights, nsim=999)
MC_cafo20
plot(MC_cafo20)
MC_cafo20$p.value

## Moran's on residuals
rm_cafo20 <- lm.morantest(m3, listw = weights, alternative = "greater")
rm_cafo20

### TB ----
## Moran's I scatterplot
NC_tb2020$lag <- lag.listw(weights, NC_tb2020$p_ir)
m4 <- lm(lag ~ p_ir, NC_tb2020)
plot(lag ~ p_ir, NC_tb2020); abline(m4, col="blue"); abline(v=mean(NC_tb2020$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2020$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb20 <- tibble(
  res = m4$residuals,
  fitted = m4$fitted.values
)

ggplot(data = r_tb20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_tb20 <- moran(NC_tb2020$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_tb20

## statistical significance
MC_tb20 <- moran.mc(NC_tb2020$p_ir, weights, nsim=999)
MC_tb20
plot(MC_tb20)
MC_tb20$p.value

## Moran's on residuals
rm_tb20 <- lm.morantest(m4, listw = weights, alternative = "greater")
rm_tb20

## Lagrange tests for TB
lm.LMtests(m4, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m4, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2023}
### CAFOs ----
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$cafo_psm)
m5 <- lm(lag ~ cafo_psm, NC_tb2023)
plot(lag ~ cafo_psm, NC_tb2023); abline(m5, col="blue"); abline(v=mean(NC_tb2023$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_tb2023$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo23 <- tibble(
  res = m5$residuals,
  fitted = m5$fitted.values
)

ggplot(data = r_cafo23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo23 <- moran(NC_tb2023$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo23

## statistical significance
MC_cafo23 <- moran.mc(NC_tb2023$cafo_psm, weights, nsim=999)
MC_cafo23
plot(MC_cafo23)
MC_cafo23$p.value

## Moran's on residuals
rm_cafo23 <- lm.morantest(m5, listw = weights, alternative = "greater")
rm_cafo23

### TB ----
## Moran's I scatterplot
NC_tb2023$lag <- lag.listw(weights, NC_tb2023$p_ir)
m6 <- lm(lag ~ p_ir, NC_tb2023)
plot(lag ~ p_ir, NC_tb2023); abline(m6, col="blue"); abline(v=mean(NC_tb2023$p_ir), lty=3, col="grey70"); abline(h=mean(NC_tb2023$p_ir), lty=3, col="grey70")

### plotting residuals
r_tb23 <- tibble(
  res = m6$residuals,
  fitted = m6$fitted.values
)

ggplot(data = r_tb23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_tb23 <- moran(NC_tb2023$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_tb23

## statistical significance
MC_tb23 <- moran.mc(NC_tb2023$p_ir, weights, nsim=999)
MC_tb23
plot(MC_tb23)
MC_tb23$p.value

## Moran's on residuals
rm_tb23 <- lm.morantest(m6, listw = weights, alternative = "greater")
rm_tb23

## Lagrange tests for TB
lm.LMtests(m6, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m6, weights, test = c("RLMerr", "RLMlag"))
```

```{r pooled}
## reorganize data
NC_tbpool <- NC_tb %>%
  group_by(County) %>%
  summarize(
    mean_ir = mean(p_ir),
    mean_cafo = mean(cafo_psm),
    mean_popdens = mean(pop_dens),
    mean_pov = mean(perc_pov),
    mean_unins = mean(perc_unins)
  )

MC_tbpool <- moran.mc(NC_tbpool$mean_ir, listw = weights, nsim = 999)
MC_tbpool
```

The CAFOs across the three years exhibit higher I than TB, which is below 0.2 for all three years but varies across them. 2020 has the lowest clustering of TB.

There's also a pretty distinct clustering of residuals and lack of constant variance with CAFOs and TB.

#### Bivariate

```{r bi Moran's}
## 2017
## get the coefficient
bm_TB2017 <- moran.bi(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights)
bm_TB2017

## statistical significance
MC <- moranbi.mc(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
## get the coefficient
bm_TB2020 <- moran.bi(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights)
bm_TB2020

## statistical significance
MC <- moranbi.mc(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
## get the coefficient
bm_TB2023 <- moran.bi(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights)
bm_TB2023

## statistical significance
MC <- moranbi.mc(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## pooled
## get the coefficient
bm_TBpool <- moran.bi(NC_tbpool$mean_ir, NC_tbpool$mean_cafo, weights)
bm_TBpool

## statistical significance
MC <- moranbi.mc(NC_tbpool$mean_ir, NC_tbpool$mean_cafo, listw = weights, nsim = 999)
MC
```

Like with the individual Moran's I, 2020 has the lowest I.

Let's also try Lee's L.

```{r Lee's L}
## 2017
l_TB2017 <- lee.test(NC_tb2017$p_ir, NC_tb2017$cafo_psm, weights)
l_TB2017

## 2020
l_TB2020 <- lee.test(NC_tb2020$p_ir, NC_tb2020$cafo_psm, weights)
l_TB2020

## 2023
l_TB2023 <- lee.test(NC_tb2023$p_ir, NC_tb2023$cafo_psm, weights)
l_TB2023
```


#### Local Moran's I

```{r local}
## 2017 ----
local_m2 <- localmoran(df_2017tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2017 <- order(df_2017tb$County)
local_m2
printCoefmat(data.frame(local_m2[county_2017,], row.names=df_2017tb$County[county_2017]), check.names=FALSE)

localg_m2 <- localG(df_2017tb$p_ir, listw = weights)
localg_m2
NC_tb2017$GetisOrd <- localg_m2
NC_tb2017 <- NC_tb2017 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2017) +
  geom_sf(aes(fill = Gi_cat))

## 2020 ----
local_m4 <- localmoran(df_2020tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2020 <- order(df_2020tb$County)
local_m4
printCoefmat(data.frame(local_m4[county_2020,], row.names=df_2020tb$County[county_2020]), check.names=FALSE)

localg_m4 <- localG(df_2020tb$p_ir, listw = weights)
localg_m4
NC_tb2020$GetisOrd <- localg_m4
NC_tb2020 <- NC_tb2020 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2020) +
  geom_sf(aes(fill = Gi_cat))

## 2023 ----
local_m6 <- localmoran(df_2023tb$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2023 <- order(df_2023tb$County)
local_m6
printCoefmat(data.frame(local_m6[county_2023,], row.names=df_2023tb$County[county_2023]), check.names=FALSE)

localg_m6 <- localG(df_2023tb$p_ir, listw = weights)
localg_m6
NC_tb2023$GetisOrd <- localg_m6
NC_tb2023 <- NC_tb2023 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_tb2023) +
  geom_sf(aes(fill = Gi_cat))
```

#### Prepare panel data

```{r prepare panel data}
df_tb <- df_tb %>%
  select(County, year, type, cat_n, County.Pop.Adjusted, p_ir, count, cafo_psm, perc_unins, perc_pov, n_foreign, Total.Land.Area, pop_dens, perc_foreign) %>%
  arrange(County, year)

panel_tb <- pdata.frame(df_tb, index = c("County", "year"))
```

#### Spatial regression

We're doing a spatial error model, which accounts for autocorrelated residuals.

```{r SEM}
sm1 <- spml(
  p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = panel_tb,
  listw = weights,
  model = "within",      # fixed effect
  effect = "twoway",     # both time and unit heterogeneity
  spatial.error = "b"
)
summary(sm1)

sm2 <- spml(
  p_ir ~ cafo_psm + pop_dens + perc_unins + perc_pov,
  data = panel_tb,
  listw = weights,
  model = "random",      # random effect
  spatial.error = "b"
)
summary(sm2)
```

We use Hausman tests to check whether the coefficients in the fixed vs random effect models differ significantly. If significant, use the FE model, because it captures group-level bias and omitted variables better, and assumes correlation between unobserved and observed characteristics

```{r Hausman}
sphtest(sm1, sm2)
```

The Hausman test tells us that the fixed effects model is the right way to go. The significance we see in the random effects model is likely due to perceived between-county associations which are actually just spatially clustered variables.

Nothing being significant is a little weird, though. It's possible that there is too little within-county, per-year variation in some of our variables (aka we have unexpectedly time-invariant predictors). Let's check that.

#### Troubleshooting

Let's compare within-county and between-county variation. Trying coefficient of variation to look at within-county variation (aka variation over time).

```{r eyeball}
summary(panel_tb)

## TB
ts_tb <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(p_ir),
            mean = mean(p_ir),
            cv = within_sd/mean)
ts_tb$cv <- ifelse(is.na(ts_tb$cv), 0, ts_tb$cv)
hist(ts_tb$cv)

## CAFO count
ts_cc <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(count),
            mean = mean(count),
            cv = within_sd/mean)
ts_cc$cv <- ifelse(is.na(ts_cc$cv), 0, ts_cc$cv)
hist(ts_cc$cv)

## CAFO density
ts_cafo <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(cafo_psm),
            mean = mean(cafo_psm),
            cv = within_sd/mean)
ts_cafo$cv <- ifelse(is.na(ts_cafo$cv), 0, ts_cafo$cv)
hist(ts_cafo$cv)

## population density
ts_pd <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(pop_dens),
            mean = mean(pop_dens),
            cv = within_sd/mean)
ts_pd$cv <- ifelse(is.na(ts_pd$cv), 0, ts_pd$cv)
hist(ts_pd$cv)

## percent poverty
ts_pov <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(perc_pov),
            mean = mean(perc_pov),
            cv = within_sd/mean)
ts_pov$cv <- ifelse(is.na(ts_pov$cv), 0, ts_pov$cv)
hist(ts_pov$cv)

## percent poverty
ts_unins <- df_tb %>% 
  group_by(County) %>%
  summarize(within_sd = sd(perc_unins),
            mean = mean(perc_unins),
            cv = within_sd/mean)
ts_unins$cv <- ifelse(is.na(ts_unins$cv), 0, ts_unins$cv)
hist(ts_unins$cv)
```

It looks like there's low within-county variation/change over time for most of the variables, so there really isn't much variation for us to model.

Another approach:

```{r variances again}
## TB
checkv_tb <- df_tb %>%
  group_by(County) %>%
  mutate(
    tb_mean = mean(p_ir, na.rm = TRUE),  # county TB incidence means
    tb_wi_resid = p_ir - tb_mean         # within-county residuals
  )

tb_total_var <- var(checkv_tb$p_ir, na.rm = TRUE)
tb_btwn_var <- var(checkv_tb$tb_mean, na.rm = TRUE); tb_btwn_var/tb_total_var
tb_wi_var <- mean(tapply(checkv_tb$tb_wi_resid, checkv_tb$County, var), na.rm = TRUE); tb_wi_var/tb_total_var

## ORDs
checkv_ord <- df_ord %>%
  group_by(County) %>%
  mutate(
    ord_mean = mean(p_ir, na.rm = TRUE),  # county ORD density means
    ord_wi_resid = p_ir - ord_mean        # within-county residuals
  )

ord_total_var <- var(checkv_ord$p_ir, na.rm = TRUE)
ord_btwn_var <- var(checkv_ord$ord_mean, na.rm = TRUE)
ord_wi_var <- mean(tapply(checkv_ord$ord_wi_resid, checkv_ord$County, var), na.rm = TRUE)
ord_btwn_var/ord_total_var

## population density
checkv_popdens <- df_tb %>%
  group_by(County) %>%
  mutate(
    popdens_mean = mean(pop_dens, na.rm = TRUE),     # county population density means
    popdens_wi_resid = pop_dens - popdens_mean       # within-county residuals
  )

popdens_total_var <- var(checkv_popdens$pop_dens, na.rm = TRUE)
popdens_btwn_var <- var(checkv_popdens$popdens_mean, na.rm = TRUE)
popdens_wi_var <- mean(tapply(checkv_popdens$popdens_wi_resid, checkv_popdens$County, var), na.rm = TRUE)
popdens_btwn_var/popdens_total_var

## poverty
checkv_pov <- df_tb %>%
  group_by(County) %>%
  mutate(
    pov_mean = mean(perc_pov, na.rm = TRUE),     # county poverty means
    pov_wi_resid = perc_pov - pov_mean           # within-county residuals
  )

pov_total_var <- var(checkv_pov$perc_pov, na.rm = TRUE)
pov_btwn_var <- var(checkv_pov$pov_mean, na.rm = TRUE)
pov_wi_var <- mean(tapply(checkv_pov$pov_wi_resid, checkv_pov$County, var), na.rm = TRUE)
pov_btwn_var/pov_total_var

## uninsured
checkv_unins <- df_tb %>%
  group_by(County) %>%
  mutate(
    unins_mean = mean(perc_unins, na.rm = TRUE),       # county uninsured means
    unins_wi_resid = perc_unins - unins_mean           # within-county residuals
  )

unins_total_var <- var(checkv_unins$perc_unins, na.rm = TRUE)
unins_btwn_var <- var(checkv_unins$unins_mean, na.rm = TRUE)
unins_wi_var <- mean(tapply(checkv_unins$unins_wi_resid, checkv_unins$County, var), na.rm = TRUE)
unins_btwn_var/unins_total_var

## cafo
checkv_cafo <- df_tb %>%
  group_by(County) %>%
  mutate(
    cafo_mean = mean(cafo_psm, na.rm = TRUE),       # county CAFO means
    cafo_wi_resid = cafo_psm - cafo_mean            # within-county residuals
  )

cafo_total_var <- var(checkv_cafo$cafo_psm, na.rm = TRUE)
cafo_btwn_var <- var(checkv_cafo$cafo_mean, na.rm = TRUE)
cafo_wi_var <- mean(tapply(checkv_cafo$cafo_wi_resid, checkv_cafo$County, var), na.rm = TRUE)
cafo_btwn_var/cafo_total_var
```

Around 99% of the total variation is due to between-county variance, meaning that there isn't enough variation for fixed-effects estimation in spatial regressions and so we got imprecise coefficient estimates.

Let's try a different approach.

#### Pooled SDM

With so little variation within counties/by year, we can pool the values across years and just run a normal, non-temporal SDM on the average values of each variable. Choosing SDM because it accounts for autocorrelation in the outcome and the predictors, as well as in the residuals. I'm also going to run an individual SDM for each year just to see how associations change over time.

I briefly ran pooled SEM and SAR models as well and the SDM beat them out with AICs.

```{r pooled SEM}
## quick map
ggplot(NC_tbpool) +
  geom_sf(aes(fill = mean_ir))

## SDM - pooled
pm1 <- lagsarlm(
  mean_ir ~ mean_cafo + mean_popdens + mean_pov + mean_unins,
  data = NC_tbpool,
  listw = weights,
  Durbin = TRUE
)
summary(pm1)
i_pm1 <- impacts(pm1, listw = weights, R = 500)
summary(i_pm1, zstats = TRUE)
MC_pm1 <- moran.mc(pm1$residuals, weights, nsim=999)
MC_pm1

## SDM by years
pm17 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2017,
  listw = weights,
  Durbin = TRUE
)
summary(pm17)
i_pm17 <- impacts(pm17, listw = weights, R = 500)
summary(i_pm17, zstats = TRUE)
MC_pm17 <- moran.mc(pm17$residuals, weights, nsim=999)
MC_pm17

pm18 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2018,
  listw = weights,
  Durbin = TRUE
)
summary(pm18)
i_pm18 <- impacts(pm18, listw = weights, R = 500)
summary(i_pm18, zstats = TRUE)
MC_pm18 <- moran.mc(pm18$residuals, weights, nsim=999)
MC_pm18

pm19 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2019,
  listw = weights,
  Durbin = TRUE
)
summary(pm19)
i_pm19 <- impacts(pm19, listw = weights, R = 500)
summary(i_pm19, zstats = TRUE)
MC_pm19 <- moran.mc(pm19$residuals, weights, nsim=999)
MC_pm19

pm20 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2020,
  listw = weights,
  Durbin = TRUE
)
summary(pm20)
i_pm20 <- impacts(pm20, listw = weights, R = 500)
summary(i_pm20, zstats = TRUE)
MC_pm20 <- moran.mc(pm20$residuals, weights, nsim=999)
MC_pm20

pm21 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2021,
  listw = weights,
  Durbin = TRUE
)
summary(pm21)
i_pm21 <- impacts(pm21, listw = weights, R = 500)
summary(i_pm21, zstats = TRUE)
MC_pm21 <- moran.mc(pm21$residuals, weights, nsim=999)
MC_pm21

pm22 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2022,
  listw = weights,
  Durbin = TRUE
)
summary(pm22)
i_pm22 <- impacts(pm22, listw = weights, R = 500)
summary(i_pm22, zstats = TRUE)
MC_pm22 <- moran.mc(pm22$residuals, weights, nsim=999)
MC_pm22

pm23 <- lagsarlm(
  p_ir ~ cafo_psm + pop_dens + perc_pov + perc_unins,
  data = NC_tb2023,
  listw = weights,
  Durbin = TRUE
)
summary(pm23)
i_pm23 <- impacts(pm23, listw = weights, R = 500)
summary(i_pm23, zstats = TRUE)
MC_pm23 <- moran.mc(pm1$residuals, weights, nsim=999)
MC_pm23
```

The nonsignificant rho and LM tests tell us that the SDMs are explaining residual autocorrelation and capturing the spatial effects, and also telling us that spatial relationships likely explained a lot of the observed CAFO-TB relationships. Good to know.

### 2. ORDs

#### Defining neighbors and making spatial weights matrix

```{r define neighbors}
nb <- poly2nb(NC_sf, queen=TRUE)
weights <- nb2listw(nb, style="W", zero.policy=TRUE)
```

#### Univariate Moran's I

We will do these for ORD (outcome) and CAFOs (key predictor). Start by getting the lags and residuals to plot and check for constant variance. Then get the global Moran's I and its p-value. Then get the Moran's I for the residuals. Lastly, calculate local Moran's.

```{r year 2017}
### CAFOs ----
NC_ord2017$lag <- lag.listw(weights, NC_ord2017$cafo_psm)
m1 <- lm(lag ~ cafo_psm, NC_ord2017)
plot(lag ~ cafo_psm, NC_ord2017); abline(m1, col="blue"); abline(v=mean(NC_ord2017$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2017$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo17 <- tibble(
  res = m1$residuals,
  fitted = m1$fitted.values
)

ggplot(data = r_cafo17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo17 <- moran(NC_ord2017$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo17

## statistical significance
MC_cafo17 <- moran.mc(NC_ord2017$cafo_psm, weights, nsim=999)
MC_cafo17
plot(MC_cafo17)
MC_cafo17$p.value

## Moran's on residuals
rm_cafo17 <- lm.morantest(m1, listw = weights, alternative = "greater")
rm_cafo17

### ORD ----
## Moran's I scatterplot
NC_ord2017$lag <- lag.listw(weights, NC_ord2017$p_ir)
m2 <- lm(lag ~ p_ir, NC_ord2017)
plot(lag ~ p_ir, NC_ord2017); abline(m2, col="blue"); abline(v=mean(NC_ord2017$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2017$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord17 <- tibble(
  res = m2$residuals,
  fitted = m2$fitted.values
)

ggplot(data = r_ord17, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_p_ir <- moran(NC_ord2017$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_p_ir

## statistical significance
MC <- moran.mc(NC_ord2017$p_ir, weights, nsim=999)
MC
plot(MC)
MC$p.value

## Moran's on residuals
rm_ord17 <- lm.morantest(m2, listw = weights, alternative = "greater")
rm_ord17

## Lagrange tests for ORD
lm.LMtests(m2, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m2, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2020}
### CAFOs ----
NC_ord2020$lag <- lag.listw(weights, NC_ord2020$cafo_psm)
m3 <- lm(lag ~ cafo_psm, NC_ord2020)
plot(lag ~ cafo_psm, NC_ord2020); abline(m3, col="blue"); abline(v=mean(NC_ord2020$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2020$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo20 <- tibble(
  res = m3$residuals,
  fitted = m3$fitted.values
)

ggplot(data = r_cafo20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo20 <- moran(NC_ord2020$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo20

## statistical significance
MC_cafo20 <- moran.mc(NC_ord2020$cafo_psm, weights, nsim=999)
MC_cafo20
plot(MC_cafo20)
MC_cafo20$p.value

## Moran's on residuals
rm_cafo20 <- lm.morantest(m3, listw = weights, alternative = "greater")
rm_cafo20

### ORD ----
## Moran's I scatterplot
NC_ord2020$lag <- lag.listw(weights, NC_ord2020$p_ir)
m4 <- lm(lag ~ p_ir, NC_ord2020)
plot(lag ~ p_ir, NC_ord2020); abline(m4, col="blue"); abline(v=mean(NC_ord2020$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2020$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord20 <- tibble(
  res = m4$residuals,
  fitted = m4$fitted.values
)

ggplot(data = r_ord20, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_ord20 <- moran(NC_ord2020$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_ord20

## statistical significance
MC_ord20 <- moran.mc(NC_ord2020$p_ir, weights, nsim=999)
MC_ord20
plot(MC_ord20)
MC_ord20$p.value

## Moran's on residuals
rm_ord20 <- lm.morantest(m4, listw = weights, alternative = "greater")
rm_ord20

## Lagrange tests for ORD
lm.LMtests(m4, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m4, weights, test = c("RLMerr", "RLMlag"))
```

```{r year 2023}
### CAFOs ----
NC_ord2023$lag <- lag.listw(weights, NC_ord2023$cafo_psm)
m5 <- lm(lag ~ cafo_psm, NC_ord2023)
plot(lag ~ cafo_psm, NC_ord2023); abline(m5, col="blue"); abline(v=mean(NC_ord2023$cafo_psm), lty=3, col="grey70"); abline(h=mean(NC_ord2023$cafo_psm), lty=3, col="grey70")

### plotting residuals
r_cafo23 <- tibble(
  res = m5$residuals,
  fitted = m5$fitted.values
)

ggplot(data = r_cafo23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_cafo23 <- moran(NC_ord2023$cafo_psm, listw=weights, n=length(nb), S0=Szero(weights))
um_cafo23

## statistical significance
MC_cafo23 <- moran.mc(NC_ord2023$cafo_psm, weights, nsim=999)
MC_cafo23
plot(MC_cafo23)
MC_cafo23$p.value

## Moran's on residuals
rm_cafo23 <- lm.morantest(m5, listw = weights, alternative = "greater")
rm_cafo23

### ORD ----
## Moran's I scatterplot
NC_ord2023$lag <- lag.listw(weights, NC_ord2023$p_ir)
m6 <- lm(lag ~ p_ir, NC_ord2023)
plot(lag ~ p_ir, NC_ord2023); abline(m6, col="blue"); abline(v=mean(NC_ord2023$p_ir), lty=3, col="grey70"); abline(h=mean(NC_ord2023$p_ir), lty=3, col="grey70")

### plotting residuals
r_ord23 <- tibble(
  res = m6$residuals,
  fitted = m6$fitted.values
)

ggplot(data = r_ord23, aes(x = fitted, y = res)) + 
  geom_point() + 
  labs(x = "Fitted values", y = "Residuals") +
  geom_hline(yintercept = 0, color = "red")

## get the coefficient
um_ord23 <- moran(NC_ord2023$p_ir, listw=weights, n=length(nb), S0=Szero(weights))
um_ord23

## statistical significance
MC_ord23 <- moran.mc(NC_ord2023$p_ir, weights, nsim=999)
MC_ord23
plot(MC_ord23)
MC_ord23$p.value

## Moran's on residuals
rm_ord23 <- lm.morantest(m6, listw = weights, alternative = "greater")
rm_ord23

## Lagrange tests for ORD
lm.LMtests(m6, weights, test = c("LMerr", "LMlag"))
lm.LMtests(m6, weights, test = c("RLMerr", "RLMlag"))
```

```{r pooled}
NC_ordpool <- NC_ord %>%
  group_by(County) %>%
  summarize(
    mean_ir = mean(p_ir),
    mean_cafo = mean(cafo_psm),
    mean_popdens = mean(pop_dens),
    mean_pov = mean(perc_pov),
    mean_unins = mean(perc_unins)
  )

MC_ordpool <- moran.mc(NC_ordpool$mean_ir, listw = weights, nsim = 999)
MC_ordpool
```

We see apparent clustering as well.

#### Bivariate

```{r bi Moran's}
## 2017
## get the coefficient
bm_ORD2017 <- moran.bi(NC_ord2017$p_ir, NC_ord2017$cafo_psm, weights)
bm_ORD2017

## statistical significance
MC <- moranbi.mc(NC_ord2017$p_ir, NC_ord2017$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2020
## get the coefficient
bm_ORD2020 <- moran.bi(NC_ord2020$p_ir, NC_ord2020$cafo_psm, weights)
bm_ORD2020

## statistical significance
MC <- moranbi.mc(NC_ord2020$p_ir, NC_ord2020$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## 2023
## get the coefficient
bm_ORD2023 <- moran.bi(NC_ord2023$p_ir, NC_ord2023$cafo_psm, weights)
bm_ORD2023

## statistical significance
MC <- moranbi.mc(NC_ord2023$p_ir, NC_ord2023$cafo_psm, weights, nsim=999)
MC
plot(MC)
MC$p.value

## pooled
## get the coefficient
bm_ORDpool <- moran.bi(NC_ordpool$mean_ir, NC_ordpool$mean_cafo, weights)
bm_ORDpool

## statistical significance
MC <- moranbi.mc(NC_ordpool$mean_ir, NC_ordpool$mean_cafo, listw = weights, nsim = 999)
MC
```

Very little if any bivariate autocorrelation.

#### Local Moran's I

```{r local}
## 2017 ----
local_m2 <- localmoran(df_2017ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2017 <- order(df_2017ord$County)
local_m2
printCoefmat(data.frame(local_m2[county_2017,], row.names=df_2017ord$County[county_2017]), check.names=FALSE)

localg_m2 <- localG(df_2017ord$p_ir, listw = weights)
localg_m2
NC_ord2017$GetisOrd <- localg_m2
NC_ord2017 <- NC_ord2017 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2017) +
  geom_sf(aes(fill = Gi_cat))

## 2020 ----
local_m4 <- localmoran(df_2020ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2020 <- order(df_2020ord$County)
local_m4
printCoefmat(data.frame(local_m4[county_2020,], row.names=df_2020ord$County[county_2020]), check.names=FALSE)

localg_m4 <- localG(df_2020ord$p_ir, listw = weights)
localg_m4
NC_ord2020$GetisOrd <- localg_m4
NC_ord2020 <- NC_ord2020 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2020) +
  geom_sf(aes(fill = Gi_cat))

## 2023 ----
local_m6 <- localmoran(df_2023ord$p_ir, listw = weights, zero.policy = attr(weights, "zero.policy"), alternative = "two.sided")
county_2023 <- order(df_2023ord$County)
local_m6
printCoefmat(data.frame(local_m6[county_2023,], row.names=df_2023ord$County[county_2023]), check.names=FALSE)

localg_m6 <- localG(df_2023ord$p_ir, listw = weights)
localg_m6
NC_ord2023$GetisOrd <- localg_m6
NC_ord2023 <- NC_ord2023 %>%
  mutate(Gi_cat = case_when(
    GetisOrd < -2.58 ~ "Very cold spot",
    GetisOrd >= -2.58 & GetisOrd < -1.96 ~ "Cold spot",
    GetisOrd >= -1.96 & GetisOrd < 1.96 ~ "Not significant",
    GetisOrd >= 1.96 & GetisOrd < 2.58 ~ "Hot spot",
    GetisOrd >= 2.58 ~ "Very hot spot"
  ))
ggplot(NC_ord2023) +
  geom_sf(aes(fill = Gi_cat))
```

Since we don't observe bivariate spatial clustering, we will not do spatial regressions here.
