---
title: "Tables and Figures"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup
```{r load packages}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(sf)
library(spdep)
library(RColorBrewer)
library(stargazer)
library(flextable)
```

### 1. CAFO locations in NC
```{r points in NC}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp")
swine_df <- readRDS("swine_df.RDS")
df <- readRDS("df.RDS")

## convert points to sf
cafo_locs <- swine_df %>%
  select(Location.Lat.Num, Location.Long.Num) %>%
  rename(lat=Location.Lat.Num, long=Location.Long.Num)
cafo_locs <- drop_na(cafo_locs)
cafo_locs <- st_as_sf(cafo_locs, coords=c("long", "lat"), crs=4326)
cafo_locs <- st_transform(cafo_locs, st_crs(NC_sf))

ggplot() +
  geom_sf(data=NC_sf, fill="grey95") +
  geom_sf(data=cafo_locs, color="red", size=1, pch=21) +
  theme_minimal()
```

### 2. Univariate choropleths
```{r univariate choropleths}
## CAFOs
class_cafo <- cut(df$cafo_dens, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_cafo)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "CAFOs") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )

## TB
class_TB <- cut(df$TB, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_TB)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "TB") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )

## other resp
class_resp <- cut(df$resp, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_resp)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "Other respiratory diseases") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )
```


### 4. Table 1

Looking at overall characteristics in North Carolina by year. Recall that in the overall df, we have duplicates for each county-year because of separate TB and ORD rows so I'm just using the df_yeartb dfs.

```{r population density}
summary(df_2017tb$pop_dens); sd(df_2017tb$pop_dens)
summary(df_2018tb$pop_dens); sd(df_2018tb$pop_dens)
summary(df_2019tb$pop_dens); sd(df_2019tb$pop_dens)
summary(df_2020tb$pop_dens); sd(df_2020tb$pop_dens)
summary(df_2021tb$pop_dens); sd(df_2021tb$pop_dens)
summary(df_2022tb$pop_dens); sd(df_2022tb$pop_dens)
summary(df_2023tb$pop_dens); sd(df_2023tb$pop_dens)
summary(df_tb$pop_dens); sd(df_tb$pop_dens)
```

```{r poverty}
summary(df_2017tb$perc_pov); sd(df_2017tb$perc_pov)
summary(df_2018tb$perc_pov); sd(df_2018tb$perc_pov)
summary(df_2019tb$perc_pov); sd(df_2019tb$perc_pov)
summary(df_2020tb$perc_pov); sd(df_2020tb$perc_pov)
summary(df_2021tb$perc_pov); sd(df_2021tb$perc_pov)
summary(df_2022tb$perc_pov); sd(df_2022tb$perc_pov)
summary(df_2023tb$perc_pov); sd(df_2023tb$perc_pov)
summary(df_tb$perc_pov); sd(df_tb$perc_pov)
```

```{r uninsured}
summary(df_2017tb$perc_unins); sd(df_2017tb$perc_unins)
summary(df_2018tb$perc_unins); sd(df_2018tb$perc_unins)
summary(df_2019tb$perc_unins); sd(df_2019tb$perc_unins)
summary(df_2020tb$perc_unins); sd(df_2020tb$perc_unins)
summary(df_2021tb$perc_unins); sd(df_2021tb$perc_unins)
summary(df_2022tb$perc_unins); sd(df_2022tb$perc_unins)
summary(df_2023tb$perc_unins); sd(df_2023tb$perc_unins)
summary(df_tb$perc_unins); sd(df_tb$perc_unins)
```

```{r foreign born}
summary(df_2017tb$perc_foreign); sd(df_2017tb$perc_foreign)
summary(df_2018tb$perc_foreign); sd(df_2018tb$perc_foreign)
summary(df_2019tb$perc_foreign); sd(df_2019tb$perc_foreign)
summary(df_2020tb$perc_foreign); sd(df_2020tb$perc_foreign)
summary(df_2021tb$perc_foreign); sd(df_2021tb$perc_foreign)
summary(df_2022tb$perc_foreign); sd(df_2022tb$perc_foreign)
summary(df_2023tb$perc_foreign); sd(df_2023tb$perc_foreign)
summary(df_tb$perc_foreign); sd(df_tb$perc_foreign)
```

```{r CAFO density}
summary(df_2017tb$cafo_psm); sd(df_2017tb$cafo_psm)
summary(df_2018tb$cafo_psm); sd(df_2018tb$cafo_psm)
summary(df_2019tb$cafo_psm); sd(df_2019tb$cafo_psm)
summary(df_2020tb$cafo_psm); sd(df_2020tb$cafo_psm)
summary(df_2021tb$cafo_psm); sd(df_2021tb$cafo_psm)
summary(df_2022tb$cafo_psm); sd(df_2022tb$cafo_psm)
summary(df_2023tb$cafo_psm); sd(df_2023tb$cafo_psm)
summary(df_tb$cafo_psm); sd(df_tb$cafo_psm)
```

```{r TB}
summary(df_2017tb$p_ir); sd(df_2017$p_ir)
summary(df_2018tb$p_ir); sd(df_2018$p_ir)
summary(df_2019tb$p_ir); sd(df_2019$p_ir)
summary(df_2020tb$p_ir); sd(df_2020$p_ir)
summary(df_2021tb$p_ir); sd(df_2021$p_ir)
summary(df_2022tb$p_ir); sd(df_2022$p_ir)
summary(df_2023tb$p_ir); sd(df_2023$p_ir)
summary(df_tb$p_ir); sd(df_tb$p_ir)
```

```{r ORD}
summary(df_2017ord$p_ir); sd(df_2017ord$p_ir)
summary(df_2018ord$p_ir); sd(df_2018ord$p_ir)
summary(df_2019ord$p_ir); sd(df_2019ord$p_ir)
summary(df_2020ord$p_ir); sd(df_2020ord$p_ir)
summary(df_2021ord$p_ir); sd(df_2021ord$p_ir)
summary(df_2022ord$p_ir); sd(df_2022ord$p_ir)
summary(df_2023ord$p_ir); sd(df_2023ord$p_ir)
summary(df_ord$p_ir); sd(df_ord$p_ir)
```


### 5. Between-county variance

Figure looking at how much of the total variance is explained by between-county variance.

```{r between-county variance}
## making table
total_var <- c(tb_total_var, ord_total_var, popdens_total_var, pov_total_var, unins_total_var, foreign_total_var, cafo_total_var)
btwn_var <- c(tb_btwn_var, ord_btwn_var, popdens_btwn_var, pov_btwn_var, unins_btwn_var, foreign_btwn_var, cafo_btwn_var)

var_df <- data.frame(btwn_var, total_var)
var_df$btwn_var <- round(var_df$btwn_var, 12); var_df$total_var <- round(var_df$total_var, 12)
var_df$variable <- c("TB incidence", "ORD incidence", "population density", "poverty rate", "uninsured rate", "foreign born", "CAFO density")
var_df$perc <- var_df$btwn_var/var_df$total_var

var_df$variable <- factor(
  var_df$variable,
  levels = c("TB incidence", "ORD incidence", "population density", "poverty rate", "uninsured rate", "foreign born", "CAFO density"),
  labels = c("TB incidence", "ORD incidence", "population density", "poverty rate", "uninsured rate", "foreign born", "CAFO density")
)

## making plot
ggplot(data = var_df, aes(x = variable, y = perc)) +
  geom_col() +
  labs(
    x = "variable",
    y = "proportion"
  ) +
  theme_classic()

ggsave("fig1.jpg", width = 10, height = 6, dpi = 1200)

```



### 6. Choropleths of variables

Faceted choropleths of pooled county characteristics and incidence rates, with univariate Moran's I.

```{r disease incidences}
## TB
quantile(NC_tbpool$mean_ir)
NC_tbpool <- NC_tbpool %>%
  mutate(ir_quant = case_when(
    mean_ir == 0 ~ "0 - None",
    mean_ir > 0 & mean_ir <= 0.39 ~ "1 - 0 to 0.39",
    mean_ir > 0.39 & mean_ir <= 1.04 ~ "2 - 0.39 to 1.04",
    mean_ir > 1.04 & mean_ir <= 1.97 ~ "3 - 1.04 to 1.97",
    mean_ir > 1.97 ~ "4 - 1.97 to 10.10"
  ))

tb_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = ir_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.36, p-val < 0.001"
  )
tb_plot

## ORDs
quantile(NC_ordpool$mean_ir)
NC_ordpool <- NC_ordpool %>%
  mutate(ir_quant = case_when(
    mean_ir == 0 ~ "0 - None",
    mean_ir > 0 & mean_ir <= 6.05 ~ "1 - 0 to 6.05",
    mean_ir > 6.05 & mean_ir <= 8.72 ~ "2 - 6.05 to 8.72",
    mean_ir > 8.72 & mean_ir <= 11.01 ~ "3 - 8.72 to 11.01",
    mean_ir > 11.01 ~ "4 - 11.01 to 26.95"
  ))

ord_plot <- ggplot(NC_ordpool) +
  geom_sf(aes(fill = ir_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.29, p-val < 0.001"
  )
ord_plot
```

```{r CAFO plot}
quantile(NC_tbpool$mean_cafo)
NC_tbpool <- NC_tbpool %>%
  mutate(cafo_quant = case_when(
    mean_cafo == 0 ~ "0 - None",
    mean_cafo > 0 & mean_cafo <= 0.0018 ~ "1 - 0 to 0.0018",
    mean_cafo > 0.0018 & mean_cafo <= 0.0083 ~ "2 - 0.0018 to 0.0083",
    mean_cafo > 0.0083 & mean_cafo <= 0.023 ~ "3 - 0.0083 to 0.023",
    mean_cafo > 0.023 ~ "4 - 0.023 to 0.59"
  ))

cafo_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = cafo_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.46, p-val < 0.001"
  )
cafo_plot
```

```{r other county characteristic plots}
## poverty
quantile(NC_tbpool$mean_pov)
NC_tbpool <- NC_tbpool %>%
  mutate(pov_quant = case_when(
    mean_pov > 7.5 & mean_pov <= 13.23 ~ "1 - 7.5 to 13.23",
    mean_pov > 13.23 & mean_pov <= 15.79 ~ "2 - 13.23 to 15.79",
    mean_pov > 15.79 & mean_pov <= 20.01 ~ "3 - 15.79 to 20.01",
    mean_pov > 20.01 ~ "4 - 20.01 to 27.90"
  ))

pov_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = pov_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.35, p-val < 0.001"
  )
pov_plot

## population density
quantile(NC_tbpool$mean_popdens)
NC_tbpool <- NC_tbpool %>%
  mutate(popdens_quant = case_when(
    mean_popdens > 7.8 & mean_popdens <= 62.74 ~ "1 - 7.8 to 62.74",
    mean_popdens > 62.74 & mean_popdens <= 112.35 ~ "2 - 62.74 to 112.35",
    mean_popdens > 112.35 & mean_popdens <= 217.59 ~ "3 - 112.35 to 217.59",
    mean_popdens > 217.59 ~ "4 - 217.59 to 2143.20"
  ))

popdens_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = popdens_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.21, p-val < 0.01"
  )
popdens_plot

## uninsured
quantile(NC_tbpool$mean_unins)
NC_tbpool <- NC_tbpool %>%
  mutate(unins_quant = case_when(
    mean_unins > 6.57 & mean_unins <= 10.44 ~ "1 - 6.57 to 10.44",
    mean_unins > 10.44 & mean_unins <= 11.47 ~ "2 - 10.44 to 11.47",
    mean_unins > 11.47 & mean_unins <= 12.5 ~ "3 - 11.47 to 12.50",
    mean_unins > 12.5 ~ "4 - 12.5 to 22.11"
  ))

unins_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = unins_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.30, p-val < 0.001"
  )
unins_plot

## foreign born
quantile(NC_tbpool$mean_fb)
NC_tbpool <- NC_tbpool %>%
  mutate(foreign_quant = case_when(
    mean_fb > 0.011 & mean_fb <= 0.029 ~ "1 - 0.011 to 0.029",
    mean_fb > 0.029 & mean_fb <= 0.041 ~ "2 - 0.029 to 0.041",
    mean_fb > 0.041 & mean_fb <= 0.059 ~ "3 - 0.041 to 0.059",
    mean_fb > 0.059 ~ "4 - 0.059 to 0.16"
  ))

foreign_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = foreign_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.33, p-val < 0.001"
  )
foreign_plot
```

Now putting it together in one figure.

```{r making one figure}
chor_plots <- ggarrange(tb_plot, ord_plot, cafo_plot, popdens_plot, pov_plot, unins_plot, foreign_plot, ncol = 3, nrow = 3, labels = "auto")
chor_plots
```


### 7. Choropleth of residuals

Faceted choropleths of pooled county characteristics and incidence rates, with univariate Moran's I.

First, make the residuals dataframes. Then plot. Then combine. Same formula as above.

```{r count info}
county_info <- NC_tbpool %>%
  select(County, geometry)
```

```{r disease incidences}
## TB
r_tbpool <- cbind(county_info, r_tbpool)
quantile(r_tbpool$res)
r_tbpool <- r_tbpool %>%
  mutate(quant = case_when(
    res <= -0.52 ~ "1 - -1.46 to -0.52",
    res > -0.52 & res <= -0.075 ~ "2 - -0.52 to -0.075",
    res > -0.075 & res <= 0.34 ~ "3 - -0.075 to 0.34",
    res > 0.34 ~ "4 - 0.34 to 2.90"
  ))

rtb_plot <- ggplot(r_tbpool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.20, p-val < 0.001"
  )
rtb_plot

## ORD
r_ordpool <- cbind(county_info, r_ordpool)
quantile(r_ordpool$res)
r_ordpool <- r_ordpool %>%
  mutate(quant = case_when(
    res <= -1.90 ~ "1 - -6.41 to -1.90",
    res > -1.90 & res <= -0.12 ~ "2 - -1.90 to -0.12",
    res > -0.12 & res <= 1.30 ~ "3 - -0.12 to 1.30",
    res > 1.30 ~ "4 - 1.30 to 8.11"
  ))

rord_plot <- ggplot(r_ordpool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.25, p-val < 0.001"
  )
rord_plot
```

```{r CAFO residuals}
r_cafopool <- cbind(county_info, r_cafopool)
quantile(r_cafopool$res)
r_cafopool <- r_cafopool %>%
  mutate(quant = case_when(
    res <= -0.017 ~ "1 - -0.099 to -0.017",
    res > -0.017 & res <= -0.012 ~ "2 - -0.017 to -0.012",
    res > -0.012 & res <= -0.0031 ~ "3 - -0.012 to -0.0031",
    res > -0.0031 ~ "4 - -0.0031 to 0.15"
  ))

rcafo_plot <- ggplot(r_cafopool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.18, p-val < 0.001"
  )
rcafo_plot
```

```{r other county characteristics}
## population density
r_popdenspool <- cbind(county_info, r_popdenspool)
quantile(r_popdenspool$res)
r_popdenspool <- r_popdenspool %>%
  mutate(quant = case_when(
    res <= -104.51 ~ "1 - -297.85 to -104.51",
    res > -104.51 & res <= -38.26 ~ "2 - -104.51 to -38.26",
    res > -38.26 & res <= 48.44 ~ "3 - -38.26 to 48.44",
    res > 48.44 ~ "4 - 48.44 to 583.08"
  ))

rpopdens_plot <- ggplot(r_popdenspool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.25, p-val < 0.001"
  )
rpopdens_plot

## poverty
r_povpool <- cbind(county_info, r_povpool)
quantile(r_povpool$res)
r_povpool <- r_povpool %>%
  mutate(quant = case_when(
    res <= -1.61 ~ "1 - -4.26 to -1.61",
    res > -1.61 & res <= -0.26 ~ "2 - -1.61 to -0.26",
    res > -0.26 & res <= 1.50 ~ "3 - -0.26 to 1.50",
    res > 1.50 ~ "4 - 1.50 to 5.35"
  ))

rpov_plot <- ggplot(r_povpool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.12, p-val < 0.05"
  )
rpov_plot

## uninsured
r_uninspool <- cbind(county_info, r_uninspool)
quantile(r_uninspool$res)
r_uninspool <- r_uninspool %>%
  mutate(quant = case_when(
    res <= -0.73 ~ "1 - -2.78 to -0.73",
    res > -0.73 & res <= -0.28 ~ "2 - -0.73 to -0.28",
    res > -0.28 & res <= 0.37 ~ "3 - -0.28 to 0.37",
    res > 0.37 ~ "4 - 0.37 to 4.32"
  ))

runins_plot <- ggplot(r_uninspool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.24, p-val < 0.001"
  )
runins_plot

## foreign born
r_fbpool <- cbind(county_info, r_fbpool)
quantile(r_fbpool$res)
r_fbpool <- r_fbpool %>%
  mutate(quant = case_when(
    res <= -0.01 ~ "1 - -0.031 to -0.010",
    res > -0.01 & res <= -0.0029 ~ "2 - -0.01 to -0.0029",
    res > -0.0029 & res <= 0.0093 ~ "3 - -0.0029 to 0.0093",
    res > 0.0093 ~ "4 - 0.0093 to 0.042"
  ))

rforeign_plot <- ggplot(r_fbpool) +
  geom_sf(aes(fill = quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    fill = "Quantile",
    subtitle = "I = 0.29, p-val < 0.001"
  )
rforeign_plot
```

```{r making one figure}
resid_plots <- ggarrange(rtb_plot, rord_plot, rcafo_plot, rpopdens_plot, rpov_plot, runins_plot, rforeign_plot, ncol = 3, nrow = 3, labels = "auto")
resid_plots
```


### 8. Lee's L scatterplot

Aka, a scatterplot of the spatially lagged x and y values.

```{r scatterplot}

```




