---
title: "Tables and Figures"
author: "Melody Xiao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 0. Setup
```{r load packages}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(sf)
library(spdep)
library(RColorBrewer)
library(stargazer)
library(flextable)
```

### 1. CAFO locations in NC
```{r points in NC}
NC_sf <- read_sf("C:/Users/mxiao/OneDrive - Duke University/Misc/CAFOs in NC/NC State County Boundary/NC_State_County_Boundary_NAD832011.shp")
swine_df <- readRDS("swine_df.RDS")
df <- readRDS("df.RDS")

## convert points to sf
cafo_locs <- swine_df %>%
  select(Location.Lat.Num, Location.Long.Num) %>%
  rename(lat=Location.Lat.Num, long=Location.Long.Num)
cafo_locs <- drop_na(cafo_locs)
cafo_locs <- st_as_sf(cafo_locs, coords=c("long", "lat"), crs=4326)
cafo_locs <- st_transform(cafo_locs, st_crs(NC_sf))

ggplot() +
  geom_sf(data=NC_sf, fill="grey95") +
  geom_sf(data=cafo_locs, color="red", size=1, pch=21) +
  theme_minimal()
```

### 2. Univariate choropleths
```{r univariate choropleths}
## CAFOs
class_cafo <- cut(df$cafo_dens, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_cafo)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "CAFOs") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )

## TB
class_TB <- cut(df$TB, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_TB)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "TB") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )

## other resp
class_resp <- cut(df$resp, 5)
ggplot(NC_sf) +
  geom_sf(aes(fill=class_resp)) +
  scale_fill_brewer(palette="OrRd") +
  labs(title = "Other respiratory diseases") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  )
```


### 4. Table 1

Looking at overall characteristics in North Carolina by year. Recall that in the overall df, we have duplicates for each county-year because of separate TB and ORD rows so I'm just using the df_yeartb dfs.

```{r population density}
summary(df_2017tb$pop_dens)
summary(df_2018tb$pop_dens)
summary(df_2019tb$pop_dens)
summary(df_2020tb$pop_dens)
summary(df_2021tb$pop_dens)
summary(df_2022tb$pop_dens)
summary(df_2023tb$pop_dens)
summary(df_tb$pop_dens)
```

```{r poverty}
summary(df_2017tb$perc_pov)
summary(df_2018tb$perc_pov)
summary(df_2019tb$perc_pov)
summary(df_2020tb$perc_pov)
summary(df_2021tb$perc_pov)
summary(df_2022tb$perc_pov)
summary(df_2023tb$perc_pov)
summary(df_tb$perc_pov)
```

```{r uninsured}
summary(df_2017tb$perc_unins)
summary(df_2018tb$perc_unins)
summary(df_2019tb$perc_unins)
summary(df_2020tb$perc_unins)
summary(df_2021tb$perc_unins)
summary(df_2022tb$perc_unins)
summary(df_2023tb$perc_unins)
summary(df_tb$perc_unins)
```

```{r CAFO density}
summary(df_2017tb$cafo_psm)
summary(df_2018tb$cafo_psm)
summary(df_2019tb$cafo_psm)
summary(df_2020tb$cafo_psm)
summary(df_2021tb$cafo_psm)
summary(df_2022tb$cafo_psm)
summary(df_2023tb$cafo_psm)
summary(df_tb$cafo_psm)
```

```{r TB}
summary(df_2017tb$p_ir)
summary(df_2018tb$p_ir)
summary(df_2019tb$p_ir)
summary(df_2020tb$p_ir)
summary(df_2021tb$p_ir)
summary(df_2022tb$p_ir)
summary(df_2023tb$p_ir)
summary(df_tb$p_ir)
```

```{r ORD}
summary(df_2017ord$p_ir)
summary(df_2018ord$p_ir)
summary(df_2019ord$p_ir)
summary(df_2020ord$p_ir)
summary(df_2021ord$p_ir)
summary(df_2022ord$p_ir)
summary(df_2023ord$p_ir)
summary(df_ord$p_ir)
```


### 5. Choropleths

Faceted choropleths of pooled county characteristics and incidence rates, with univariate Moran's I and Lee's L values.

```{r disease incidences}
## TB
quantile(NC_tbpool$mean_ir)
NC_tbpool <- NC_tbpool %>%
  mutate(ir_quant = case_when(
    mean_ir == 0 ~ "0 - None",
    mean_ir > 0 & mean_ir <= 0.39 ~ "1 - 0 to 0.39",
    mean_ir > 0.39 & mean_ir <= 1.04 ~ "2 - 0.39 to 1.04",
    mean_ir > 1.04 & mean_ir <= 1.97 ~ "3 - 1.04 to 1.97",
    mean_ir > 1.97 ~ "4 - over 1.97"
  ))

tb_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = ir_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "TB incidence by county",
    subtitle = "Cases per 100,000",
    fill = "Quantile"
  )
tb_plot

## ORDs
quantile(NC_ordpool$mean_ir)
NC_ordpool <- NC_ordpool %>%
  mutate(ir_quant = case_when(
    mean_ir == 0 ~ "0 - None",
    mean_ir > 0 & mean_ir <= 6.05 ~ "1 - 0 to 6.05",
    mean_ir > 6.05 & mean_ir <= 8.72 ~ "2 - 6.05 to 8.72",
    mean_ir > 8.72 & mean_ir <= 11.01 ~ "3 - 8.72 to 11.01",
    mean_ir > 11.01 ~ "4 - over 11.01"
  ))

ord_plot <- ggplot(NC_ordpool) +
  geom_sf(aes(fill = ir_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "ORD incidence by county",
    subtitle = "Cases per 100,000",
    fill = "Quantile"
  )
ord_plot
```

```{r CAFO plot}
quantile(NC_tbpool$mean_cafo)
NC_tbpool <- NC_tbpool %>%
  mutate(cafo_quant = case_when(
    mean_cafo == 0 ~ "0 - None",
    mean_cafo > 0 & mean_cafo <= 0.0018 ~ "1 - 0 to 0.0018",
    mean_cafo > 0.0018 & mean_cafo <= 0.0083 ~ "2 - 0.0018 to 0.0083",
    mean_cafo > 0.0083 & mean_cafo <= 0.023 ~ "3 - 0.0083 to 0.023",
    mean_cafo > 0.023 ~ "4 - over 0.023"
  ))

cafo_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = cafo_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "CAFO density by county",
    subtitle = "Facilities per square mile",
    fill = "Quantile"
  )
cafo_plot
```

```{r other county characteristic plots}
## poverty
quantile(NC_tbpool$mean_pov)
NC_tbpool <- NC_tbpool %>%
  mutate(pov_quant = case_when(
    mean_pov > 7.5 & mean_pov <= 13.23 ~ "1 - 7.5 to 13.23",
    mean_pov > 13.23 & mean_pov <= 15.79 ~ "2 - 13.23 to 15.79",
    mean_pov > 15.79 & mean_pov <= 20.01 ~ "3 - 15.79 to 20.01",
    mean_pov > 20.01 ~ "4 - over 20.01"
  ))

pov_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = pov_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Poverty rate by county",
    subtitle = "Percent living under federal poverty line",
    fill = "Quantile"
  )
pov_plot

## population density
quantile(NC_tbpool$mean_popdens)
NC_tbpool <- NC_tbpool %>%
  mutate(popdens_quant = case_when(
    mean_popdens > 7.8 & mean_popdens <= 62.74 ~ "1 - 7.8 to 62.74",
    mean_popdens > 62.74 & mean_popdens <= 112.35 ~ "2 - 62.74 to 112.35",
    mean_popdens > 112.35 & mean_popdens <= 217.59 ~ "3 - 112.35 to 217.59",
    mean_popdens > 217.59 ~ "4 - over 217.59"
  ))

popdens_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = popdens_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Population density by county",
    subtitle = "People per square mile",
    fill = "Quantile"
  )
popdens_plot

## uninsured
quantile(NC_tbpool$mean_unins)
NC_tbpool <- NC_tbpool %>%
  mutate(unins_quant = case_when(
    mean_unins > 6.57 & mean_unins <= 10.44 ~ "1 - 6.57 to 10.44",
    mean_unins > 10.44 & mean_unins <= 11.47 ~ "2 - 10.44 to 11.47",
    mean_unins > 11.47 & mean_unins <= 12.5 ~ "3 - 11.47 to 12.50",
    mean_unins > 12.5 ~ "4 - over 12.5"
  ))

unins_plot <- ggplot(NC_tbpool) +
  geom_sf(aes(fill = unins_quant)) +
  scale_fill_brewer(palette = "YlGnBu") +
  theme(
    panel.grid = element_blank(),
    panel.background = element_blank(),
    plot.background = element_blank(),
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank()
  ) +
  labs(
    title = "Uninsured rate by county",
    subtitle = "Percent uninsured",
    fill = "Quantile"
  )
unins_plot
```





